---
title: "age_group_analysis"
author: 
date: "`r Sys.Date()`"
output: html_document
---
#load libraries 
```{r}
library(librarian)
shelf(dplyr)
shelf(binom)
shelf(tidyr)
shelf(ggplot2)
shelf(scales)
shelf(purrr)
shelf(tibble)
shelf(broom)
shelf(multiwayvcov)
shelf(lmtest)
shelf(pROC)
```

#MAKE AGE DATA FRAME 
```{r}
age_data_df <- all_data_df %>%
  dplyr::left_join(
    dplyr::select(kpp_data, studyno, ageyr),
    by = c("sample1" = "studyno")
  ) %>%
  dplyr::rename(age1 = ageyr) %>%
  
  dplyr::left_join(
    dplyr::select(kpp_data, studyno, ageyr),
    by = c("sample2" = "studyno")
  ) %>%
  dplyr::rename(age2 = ageyr) %>%
  
  dplyr::mutate(
    age_group = dplyr::case_when(
      age1 <  5              ~ "<5y",
      age1 >= 5 & age1 <= 10 ~ "5-10y",
      TRUE                   ~ ">10y"
    ),
    age_group = factor(age_group,
                       levels = c("<5y", "5-10y", ">10y"))
  ) 
```

#Source 2A Binning Function
```{r}
source("script/clean/helper_scripts/helper_2A_binning.R")
```

#NEW BINS WITH FIXED FOCAL CASE 
```{r}
shelf(dplyr, tidyr, purrr, ggplot2)


age_pair_df <- all_data_df %>%
  dplyr::left_join(
    dplyr::select(kpp_data, studyno, admitdate, ageyr),
    by = c("sample1" = "studyno")
  ) %>%
  dplyr::rename(admit1 = admitdate, age1 = ageyr) %>%
  dplyr::mutate(admit1 = as.Date(admit1)) %>%
  dplyr::left_join(
    dplyr::select(kpp_data, studyno, admitdate, ageyr),
    by = c("sample2" = "studyno")
  ) %>%
  dplyr::rename(admit2 = admitdate, age2 = ageyr) %>%
  dplyr::mutate(admit2 = as.Date(admit2))

# 1) Re-order so sample1 is always the earlier-admit (focal) case
ordered_age_pairs <- age_pair_df %>%
  dplyr::mutate(first1 = admit1 <= admit2) %>%
  dplyr::mutate(
    sample1 = if_else(first1, sample1, sample2),
    sample2 = if_else(first1, sample2, sample1),
    admit1  = if_else(first1, admit1,  admit2),
    admit2  = if_else(first1, admit2,  admit1),
    age1    = if_else(first1, age1,    age2),
    age2    = if_else(first1, age2,    age1)
  ) %>%
  dplyr::select(-first1)


ordered_age_pairs <- ordered_age_pairs %>%
  dplyr::mutate(
    age_group = dplyr::case_when(
      age1 <  5             ~ "<5y",
      age1 >= 5 & age1 <= 10 ~ "5-10y",
      TRUE                   ~ ">10y"
    ) %>%
    factor(levels = c("<5y","5-10y",">10y"))
  )


ref_age_binned <- ordered_age_pairs %>%
  filter(age_group == ">10y") %>%
  fc_2A_binning() %>%
  rename(
    prop_ref     = prop_same_chain,
    ci_lower_ref = lower,
    ci_upper_ref = upper
  )

age_diff_df <- ordered_age_pairs %>%
  filter(age_group != ">10y") %>%
  group_by(age_group) %>%
  group_modify(~ .x %>%
    fc_2A_binning() %>%
    rename(
      prop_grp     = prop_same_chain,
      ci_lower_grp = lower,
      ci_upper_grp = upper
    ) %>%
    left_join(ref_age_binned, by = "bin_cutoff") %>%
    mutate(
      prop_diff     = prop_grp - prop_ref,
      se_diff       = sqrt(
                        ((ci_upper_grp - prop_grp)/1.96)^2 +
                        ((ci_upper_ref - prop_ref)/1.96)^2
                      ),
      ci_lower_diff = prop_diff - 1.96 * se_diff,
      ci_upper_diff = prop_diff + 1.96 * se_diff
    )
  ) %>%
  ungroup()

ggplot(age_diff_df, aes(bin_cutoff, prop_diff, colour=age_group, linetype=age_group)) +
  geom_ribbon(aes(ymin=ci_lower_diff, ymax=ci_upper_diff, fill=age_group),
              alpha=0.3, colour=NA) +
  geom_line(size=1.2) +
  geom_hline(yintercept=0, linetype="dashed", colour="grey40") +
  scale_x_continuous("Distance (km)", limits=c(0,30), breaks=seq(0,30,5), expand=c(0,0)) +
  scale_y_continuous("Δ P(same chain) (<5y & 5-10y vs >10y)", limits=c(-0.02,0.1), expand=c(0,0)) +
  scale_colour_manual("Age group", values=c("<5y"="#1b9e77","5-10y"="#d95f02")) +
  scale_fill_manual(guide="none", values=c("<5y"="#1b9e77","5-10y"="#d95f02")) +
  scale_linetype_manual("Age group", values=c("<5y"="solid","5-10y"="dashed")) +
  theme_minimal(base_size=14) +
  theme(
    panel.grid = element_blank(),
    axis.line  = element_line(size=0.6, colour="black"),
    axis.title = element_text(face="bold")
  )
```

#glm for fixed focal case 
```{r}
ordered_age_pairs <- ordered_age_pairs %>%
  filter(!is.na(spatial_dist)) %>%
  mutate(
    age_group = factor(age_group, levels = c(">10y", "5-10y", "<5y")),
    log_dist  = log(spatial_dist + 1e-6)
  )

glm_age_log <- glm(
  same_chain ~ age_group * log_dist,
  data   = ordered_age_pairs,
  family = binomial(link = "logit")
)

vcov_cl <- cluster.vcov(
  glm_age_log,
  cluster = list(
    ordered_age_pairs$sample1,
    ordered_age_pairs$sample2
  )
)

age_or_table_log_robust <- broom::tidy(
  glm_age_log,
  conf.int    = TRUE,
  exponentiate= TRUE,
  vcov.       = vcov_cl
) %>%
  mutate(
    OR      = sprintf("%.3f", estimate),
    `95% CI`= sprintf("[%.3f–%.3f]", conf.low, conf.high),
    `P-value` = ifelse(p.value < 0.001, "<0.001", sprintf("%.3f", p.value))
  ) %>%
  select(
    Term    = term,
    `OR`    = OR,
    `95% CI`= `95% CI`,
    `P-value`
  )

age_or_table_log_robust

```

#test diagnostics 
```{r}
mf <- model.frame(glm_age_log)

response <- model.response(mf)               
pred_prob <- predict(glm_age_log, newdata = mf, type = "response")

roc_obj <- roc(response, pred_prob)
roc_obj

brier <- mean((pred_prob - response)^2)
brier
```


####ASSORTATIVITY 

####2 year Y/N analysis 
```{r}
age_data_df <- age_data_df %>%
  mutate(
    log_dist        = log(spatial_dist + 1e-6),
    age_diff        = abs(age1 - age2),
    within_2yrs_num = as.integer(age_diff <= 2),              
    within_2yrs_yn  = if_else(age_diff <= 2, "yes", "no")     
  )

age_data_df %>%
  count(within_2yrs_yn)
```
##IMPLEMENT THE NEW ASSORTATIVITY MODEL WITH BOOTSTRAP


# 1) Define the formula
#fmla <- same_chain ~ within_2yrs_num * log_dist + serotype

# 2) Fit the logistic regression
#glm_model <- glm(formula = fmla, data = age_data_df, family  = binomial(link = "logit"))

#Fit model and do bootstrap 
```{r}
shelf(dplyr)
shelf(speedglm)
shelf(tibble)

set.seed(2025)

df0 <- age_data_df %>%
  filter(!is.na(spatial_dist)) %>%
  mutate(
    log_dist        = log(spatial_dist + 1e-6),
    age_diff        = abs(age1 - age2),
    within_2yrs_num = as.integer(age_diff <= 2),
    serotype        = factor(serotype)
  )

fmla <- same_chain ~ within_2yrs_num * (log_dist + serotype)

two_way_boot <- function(data, formula, cl1, cl2, B = 500) {
  ids1 <- unique(data[[cl1]])
  ids2 <- unique(data[[cl2]])
  n1   <- length(ids1)
  n2   <- length(ids2)
  
  base_fit <- speedglm(formula, data, family=binomial(link="logit"), accel=TRUE)
  coef_names <- names(coef(base_fit))
  p          <- length(coef_names)
  
  boots <- matrix(NA, nrow = B, ncol = p, dimnames = list(NULL, coef_names))
  
  for (b in seq_len(B)) {
    draw1 <- sample(ids1, size = n1, replace = TRUE)
    draw2 <- sample(ids2, size = n2, replace = TRUE)
    
    boot_df <- data %>%
      filter( .data[[cl1]] %in% draw1,
              .data[[cl2]] %in% draw2 )
    
    fit_b <- speedglm(
      formula = formula,
      data    = boot_df,
      family  = binomial(link = "logit"),
      accel   = TRUE
    )
    
    boots[b, ] <- coef(fit_b)
  }
  
  ci_log   <- apply(boots, 2, quantile, probs = c(0.025, 0.975))
  orig_log <- coef(base_fit)
  
  tibble(
    Term     = coef_names,
    OR       = sprintf("%.3f", exp(orig_log)),
    `95% CI` = sprintf("[%.3f–%.3f]",
                       exp(ci_log[1, ]),
                       exp(ci_log[2, ]))
  )
}

reg_table <- two_way_boot(df0, fmla, "sample1", "sample2", B = 500)
print(reg_table)
```

#Final table marking significance 
```{r}
reg_table_sig <- reg_table %>%
  mutate(CI_clean = gsub("^\\[|\\]$", "", `95% CI`)) %>%
  separate(CI_clean, into = c("lo", "hi"), sep = "–", convert = TRUE) %>%
  mutate(significant = if_else(lo > 1 | hi < 1, "yes", "no")) %>%
  dplyr::select(Term, OR, `95% CI`, significant)


print(reg_table_sig)
```

```{r}
cat(
  "AUC =",
  round(
    pROC::auc(
      df0$same_chain,
      predict(
        speedglm(formula = fmla, data = df0, family = binomial(link = "logit"), accel = TRUE),
        newdata = df0,
        type    = "response"
      )
    ),
    3
  ),
  "\n"
)
```




