---
title: "combining_serotypes"
author: 
date: "`r Sys.Date()`"
output: html_document
---
#Load Libraries 
```{r}
library(librarian)
shelf(dplyr)
shelf(purrr)
shelf(tidyverse)
shelf(scales)
shelf(minpack.lm)
shelf(car)
shelf(ResourceSelection)
shelf(pROC)
shelf(Metrics)
shelf(ggplot2)
```


#Same_chain is defined in each df 
#all_data_df combines each serotype data to make one df 

#For each distance bin, it counts the total number of pairs within that cutoff and the number flagged as same_chain, computes the observed proportion of linked pairs, and then bootstraps for 95% confidence interval for that proportion

#Make Massive DF 
```{r}
all_data_df <- bind_rows(
  d1_pairs_df %>% mutate(serotype = "DENV1"),
  d2_pairs_df %>% mutate(serotype = "DENV2"),
  d3_pairs_df %>% mutate(serotype = "DENV3"),
  d4_pairs_df %>% mutate(serotype = "DENV4")
)
```

#cumulative distance bins of width 0.75 km.
#include every pair whose spatial separation < distance 
#cumulative proportion


#Source 2A Binning Function
```{r}
source("script/clean/helper_scripts/helper_2A_binning.R")
```

#Make 2A Binning Frame 
```{r}
chain_summary_df <- fc_2A_binning(all_data_df)
```

#Plot 2A
```{r}
all_serotype_prop <- ggplot(chain_summary_df, aes(x = bin_cutoff, y = prop_same_chain)) +
  geom_point(shape = 21, size = 3, stroke = 1.2, color = "blue", fill = "white") +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, alpha = 0.8, color = "blue") +
  
  scale_y_continuous(
    name   = "Proportion Same Chain",
    breaks = seq(0, 1, by = 0.01),
    labels = scales::number_format(accuracy = 0.01),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, by = 5),
    labels = scales::number_format(accuracy = 1)
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    panel.grid      = element_blank(),
    axis.line       = element_line(size = 0.6, color = "black"),
    axis.ticks      = element_line(size = 0.6, color = "black"),
    axis.title      = element_text(size = 14, face = "plain", color = "black"),
    axis.text       = element_text(size = 12, color = "black"),
    axis.text.x     = element_text(angle = 45, hjust = 1),
    plot.title      = element_blank(),
    legend.title    = element_blank(),
    legend.position = "top"
  )

all_serotype_prop


```

##CHAIN DF EXTRAPOLATION 
```{r}
df_emp <- chain_summary_df %>%
  mutate(
    d_mid = (lower + upper) / 2
  ) %>%
  filter(d_mid > 0) %>%      # drop the 0‐km bin
  arrange(d_mid)

P0 <- df_emp$prop_same_chain[1]
half_prop <- P0/2

idx <- which(df_emp$prop_same_chain <= half_prop)[1]
if (!is.na(idx) && idx > 1) {
  d1 <- df_emp$d_mid[idx-1]; p1 <- df_emp$prop_same_chain[idx-1]
  d2 <- df_emp$d_mid[idx];   p2 <- df_emp$prop_same_chain[idx]
  d50 <- d1 + (half_prop - p1)*(d2 - d1)/(p2 - p1)
} else {
  d50 <- NA
}

df_rates <- df_emp %>%
  mutate(b_i = -log(prop_same_chain / P0) / d_mid,
         d50_i = log(2)/b_i)

b_ci    <- quantile(df_rates$b_i,   c(0.025, 0.975), na.rm = TRUE)
d50_ci  <- quantile(df_rates$d50_i, c(0.025, 0.975), na.rm = TRUE)

empirical_metrics <- tibble(
  empirical_d50   = d50,
  d50_lo          = d50_ci[1],
  d50_hi          = d50_ci[2],
  median_b        = median(df_rates$b_i, na.rm = TRUE),
  b_lo            = b_ci[1],
  b_hi            = b_ci[2]
)

print(empirical_metrics)
```


#LOGISTIC REGRESSION FIT 
```{r}
mod_glm <- glm(
  same_chain ~ log(spatial_dist + 1e-6),
  family = binomial(link = "logit"),
  data   = all_data_df
)

term_name <- grep("^log\\(spatial_dist", names(coef(mod_glm)), value = TRUE)
beta        <- coef(mod_glm)[term_name]
ci_mat      <- confint(mod_glm)
ci_beta     <- ci_mat[term_name, ]
OR_log_km   <- exp(beta)
OR_CI_log   <- exp(ci_beta)
pct_drop_OR <- (1 - OR_log_km) * 100

cat(sprintf(
  "Logistic regression (per log‐km):\n  Odds ratio = %.3f (95%% CI %.3f–%.3f)\n  Approximate %.1f%% decrease in odds per log‐km\n\n",
  OR_log_km, OR_CI_log[1], OR_CI_log[2], pct_drop_OR
))
```

#Exponential Function Fit 
```{r}
decay_fit <- nlsLM(
  prop_same_chain ~ a * exp(-b * bin_cutoff),
  data  = chain_summary_df,
  start = list(a = 1, b = 0.1)
)
```


#Logistic Checks 
```{r}
bt_log <- boxTidwell(
  same_chain ~ log(spatial_dist + 1e-6),
  data = all_data_df %>% filter(spatial_dist > 0)
)
print(bt_log)  

hl_log <- hoslem.test(
  mod_glm$y,
  fitted(mod_glm),
  g = 10
)
print(hl_log)   

auc_log   <- auc(mod_glm$y, fitted(mod_glm))
brier_log <- mean((fitted(mod_glm) - mod_glm$y)^2)
cat("Log‐distance model AUC =", round(auc_log, 3), "\n")
cat("Log‐distance model Brier score =", round(brier_log, 4), "\n")

# 4) Overdispersion check
phi_log <- sum(residuals(mod_glm, type = "pearson")^2) / mod_glm$df.residual
cat("Log‐distance model dispersion φ̂ =", round(phi_log, 3), "\n")


res_decay  <- residuals(decay_fit)
pred_decay <- predict(decay_fit)
plot(pred_decay, res_decay,
     xlab = "Fitted P(d)", ylab = "Residuals",
     main = "Residuals vs Fitted: Exponential‐decay")
abline(h = 0, lty = 2)

rmse_decay <- rmse(chain_summary_df$prop_same_chain, pred_decay)
sse        <- sum((chain_summary_df$prop_same_chain - pred_decay)^2)
sst        <- sum((chain_summary_df$prop_same_chain -
                   mean(chain_summary_df$prop_same_chain))^2)
r2_decay   <- 1 - sse/sst
cat("Exponential RMSE =", round(rmse_decay, 4), "\n")
cat("Exponential pseudo‐R² =", round(r2_decay, 3), "\n")

```

##EXPONENTIAL CHECK PLS
```{r}
decay_fit <- nlsLM(
  prop_same_chain ~ a * exp(-b * bin_cutoff),
  data  = chain_summary_df,
  start = list(a = 1, b = 0.1)
)

fitted_d <- predict(decay_fit)
resid_d  <- chain_summary_df$prop_same_chain - fitted_d

rmse_d <- sqrt(mean(resid_d^2))
sst_d  <- sum((chain_summary_df$prop_same_chain - mean(chain_summary_df$prop_same_chain))^2)
sse_d  <- sum(resid_d^2)
r2_d   <- 1 - sse_d/sst_d
cat(sprintf("RMSE = %.4f\npseudo‐R² = %.3f\n\n", rmse_d, r2_d))

ggplot(data.frame(fitted = fitted_d, resid = resid_d),
       aes(x = fitted, y = resid)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Residuals vs Fitted", x = "Fitted P(d)", y = "Residuals") +
  theme_minimal()

qqnorm(resid_d, main = "Q–Q Plot of Residuals"); qqline(resid_d)
```

#3 Parameter exponential FUNCTION
```{r}
start_vals <- list(
  c = min(chain_summary_df$prop_same_chain, na.rm = TRUE),                    #
  a = max(chain_summary_df$prop_same_chain, na.rm = TRUE) -
      min(chain_summary_df$prop_same_chain, na.rm = TRUE),                    #
  b = 0.1                                                                     
)

decay_model <- nlsLM(
  prop_same_chain ~ c + a * exp(-b * bin_cutoff),
  data    = chain_summary_df,
  start   = start_vals,
  weights = 1 / (se^2)
)

summary(decay_model)

ci <- confint(decay_model, level = 0.95)
print(ci)

ci_rounded <- round(ci, 4)
colnames(ci_rounded) <- c("2.5 %", "97.5 %")
ci_rounded
```

```{r}
coef_vals     <- coef(decay_model)
plateau_c     <- coef_vals["c"]
amplitude_a   <- coef_vals["a"]
decay_rate_b  <- coef_vals["b"]
half_distance <- log(2) / decay_rate_b
initial_p0    <- plateau_c + amplitude_a

cat(sprintf(
  "Fitted parameters:\n • c (plateau)     = %.4f\n • a (amplitude)   = %.4f\n • b (decay rate)  = %.4f\n • p(0)            = %.4f\n • half-distance   = %.2f km\n",
  plateau_c, amplitude_a, decay_rate_b, initial_p0, half_distance
))
```


```{r}
ci <- confint(decay_model, level = 0.95)

b_low  <- ci["b", 1]
b_high <- ci["b", 2]

b_est  <- coef(decay_model)["b"]

# Compute half‐distance and its 95% CI
half_dist_est   <- log(2) / b_est
half_dist_lower <- log(2) / b_high   
half_dist_upper <- log(2) / b_low

# Report
cat(sprintf(
  "Half‐distance = %.2f km (95%% CI: %.2f–%.2f km)\n",
  half_dist_est, half_dist_lower, half_dist_upper
))

```



#Check 3-Param Model Fit 
```{r}
chain_summary_df <- chain_summary_df %>%
  mutate(
    fitted   = predict(decay_model),
    residual = prop_same_chain - fitted
  )

obs <- chain_summary_df$prop_same_chain
pred <- chain_summary_df$fitted
pseudo_R2 <- 1 - sum((obs - pred)^2) / sum((obs - mean(obs))^2)
cat(sprintf("Pseudo-R² = %.3f\n", pseudo_R2))

rmse <- sqrt(mean((obs - pred)^2))
cat(sprintf("RMSE       = %.4f\n", rmse))

brier <- mean((obs - pred)^2)
cat(sprintf("Brier score = %.6f\n", brier))
```

##########################################

#Source Prep Functon for 1G 
```{r}
source("script/clean/helper_scripts/helper_1G_binning.R")
```

#Prep for 1G
```{r}
cumulative_df <- fc_1G_binning(all_data_df) 

overall_median <- all_data_df %>% 
  summarise(median_spatial = median(spatial_dist, na.rm = TRUE)) %>% 
  pull(median_spatial)
```

#Plot for 1G 
```{r}
all_serotype_evol_time <- ggplot(cumulative_df, aes(x = evolutionary_time, y = spatial_mean)) +
  geom_ribbon(aes(ymin = spatial_lower, ymax = spatial_upper),
              fill = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = overall_median,
             linetype = "dashed",
             color    = "black") +
  scale_x_continuous(
    name = "Evolutionary time (years)",
    sec.axis = sec_axis(
      trans = ~ . / generation_time_years,
      name  = "Transmission generations (No.)"
    )
  ) +
  theme_minimal() +
  theme(
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top  = element_text(size = 10),
    axis.line.x.top  = element_line(color = "black")
  ) +
  labs(
    y = "Spatial distance (km)"
  )

all_serotype_evol_time

ggsave("1G_all_serotype_.pdf", plot = all_serotype_evol_time, device = "pdf", width = 8, height = 6)
```

##COMPARING TIME 

#Make Data Frame 
```{r}
period_lookup_df <- bind_rows(
  d1_94_06_pairs_df,
  d1_07_19_pairs_df,
  d2_94_06_pairs_df,
  d2_07_19_pairs_df,
  d3_94_06_pairs_df,
  d3_07_19_pairs_df,
  d4_94_06_pairs_df,
  d4_07_19_pairs_df)
```

#Source Same Chain
```{r}
source("script/clean/helper_scripts/helper_same_chain.R")
```

#Function for same chain 
```{r}
period_lookup_df <- fc_same_chain(period_lookup_df)
sum(period_lookup_df$same_chain, na.rm = TRUE)
```

#Source Function for 1G Binning 
```{r}
source("script/clean/helper_scripts/helper_1G_binning.R")
```

#Binning for Plot 1G for All Serotype Time Comparison 
```{r}
period1_df <- period_lookup_df %>% 
  filter(period == "1994-2006") %>% 
  fc_1G_binning() %>% 
  mutate(period = "1994-2006")

period2_df <- period_lookup_df %>% 
  filter(period == "2007-2019") %>% 
  fc_1G_binning() %>% 
  mutate(period = "2007-2019")

period_cumulative_df <- bind_rows(period1_df, period2_df)

overall_median        <- bind_rows(d1_pairs_df, d2_pairs_df, d3_pairs_df, d4_pairs_df) %>% 
                         pull(spatial_dist) %>% 
                         median(na.rm = TRUE)
generation_time_years <- 20 / 365
```

#Plot 1G for All Serotype Time Comparison 
```{r}
all_sero_period_evol_time <- ggplot(period_cumulative_df, 
    aes(x = evolutionary_time, y = spatial_mean, color = period, fill = period)) +
  geom_ribbon(aes(ymin = spatial_lower, ymax = spatial_upper),
              alpha = 0.4, color = NA) +
  geom_point(size = 1.5) +
  geom_line(size = 1) +
  geom_hline(yintercept = overall_median,
             linetype = "dashed",
             color    = "black") +
  scale_x_continuous(
    name = "Evolutionary time (years)",
    sec.axis = sec_axis(
      trans = ~ . / generation_time_years,
      name  = "Transmission generations (No.)"
    )
  ) +
  scale_color_manual(values = c("1994-2006" = "blue",
                                "2007-2019" = "darkred")) +
  scale_fill_manual(values = c("1994-2006" = "blue",
                               "2007-2019" = "darkred")) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top  = element_text(size = 10),
    axis.line.x.top  = element_line(color = "black"),
    legend.title     = element_blank(),
    legend.position  = "top"
  ) +
  labs(y = "Spatial distance (km)")

all_sero_period_evol_time

#ggsave("1G_all_period.pdf", plot = all_sero_period_evol_time, device = "pdf", width = 8, height = 6)
```

#Source 2A Binning Function
```{r}
source("script/clean/helper_scripts/helper_2A_binning.R")
```

#Binning for Plot 2A for All Serotype Time Comparison 
```{r}
period1_df_2A <- period_lookup_df %>%
  filter(period == "1994-2006") %>%
  fc_2A_binning() %>%
  transmute(
    period        = "1994-2006",
    distance_km   = bin_cutoff,
    total_pairs   = n,
    chain_pairs   = same_chain_count,
    prop_chain    = prop_same_chain,
    se_chain      = se,
    ci_lower      = lower,
    ci_upper      = upper
  )

period2_df_2A <- period_lookup_df %>%
  filter(period == "2007-2019") %>%
  fc_2A_binning() %>%
  transmute(
    period        = "2007-2019",
    distance_km   = bin_cutoff,
    total_pairs   = n,
    chain_pairs   = same_chain_count,
    prop_chain    = prop_same_chain,
    se_chain      = se,
    ci_lower      = lower,
    ci_upper      = upper
  )

all_chain_summary <- bind_rows(period1_df_2A, period2_df_2A) %>%
  filter(!is.na(prop_chain))
```

#Plot 2A for All Serotype Time Comparison 
```{r}
all_prop_chain_time <- ggplot(all_chain_summary,
                              aes(x = distance_km,
                                  y = prop_chain,
                                  color = period,
                                  group = period)) +
  geom_point(shape = 21, size = 3, stroke = 1.2, fill = "white") +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper),
                width = 0.2) +
  geom_line(size = 1) +
  
  scale_y_continuous(
    name   = "Proportion Same Chain",
    breaks = seq(0, 1, by = 0.02),
    labels = number_format(accuracy = 0.01),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, by = 5),
    labels = number_format(accuracy = 1)
  ) +
  
  labs(
    title = "Proportion of Pairs in Same Transmission Chain by Period"
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    panel.grid      = element_blank(),
    axis.line       = element_line(size = 0.6, color = "black"),
    axis.ticks      = element_line(size = 0.6, color = "black"),
    axis.title      = element_text(size = 14, face = "bold"),
    axis.text.x     = element_text(angle = 45, hjust = 1),
    axis.text       = element_text(size = 12),
    legend.title    = element_blank(),
    legend.position = "top",
    plot.title      = element_text(size = 16, face = "bold", hjust = 0.5)
  )

print(all_prop_chain_time)

#ggsave("2A_all_period.pdf", plot = all_prop_chain_time, device = "pdf", width = 8, height = 6)
```

