---
title: "epi_size"
author: 
date: "`r Sys.Date()`"
output: html_document
---

#Load Packages
```{r}
library(librarian)
shelf(dplyr)
shelf(ggplot2)
shelf(scales)
shelf(viridis)
shelf(broom)
shelf(ggeffects)
shelf(ggplot2)
shelf(RColorBrewer)
```

#Load Data
```{r}
kpp_data <- read.csv("data/raw/kpp_data.csv", stringsAsFactors = FALSE)
```


#Make Temporary Data Frame
```{r}
pairwise_tmp <- pairwise_df_with_periods %>%
  mutate(
    serotype = toupper(serotype),    
    year1    = as.integer(year1),
    year2    = as.integer(year2)
  )

cases_tmp <- kpp_data %>%
  filter(!is.na(DENV.PCR), !is.na(study.year)) %>%
  mutate(
    serotype = toupper(DENV.PCR),
    serotype = ifelse(grepl("^DEN[1-4]$", serotype),
                      sub("^DEN([1-4])$", "DENV\\1", serotype),
                      serotype),
    year = as.integer(study.year)
  ) %>%
  count(serotype, year, name = "cases")

print(cases_tmp)

```

#Plotting Frame
```{r}
yearly_summary <- pairwise_tmp %>%
  filter(!is.na(same_chain), year1 == year2) %>%
  group_by(serotype, year = year1) %>%
  summarize(
    prop_same_chain = mean(same_chain, na.rm = TRUE),
    n_pairs         = n(),
    .groups = "drop"
  ) %>%
  left_join(cases_tmp, by = c("serotype", "year")) 


head(yearly_summary)
```


```{r}
ggplot(yearly_summary,
       aes(x     = year,
           y     = prop_same_chain,
           size  = cases,
           color = serotype,
           group = serotype)) +
  
  geom_line(size = 0.8) +
  geom_point(shape = 16, alpha = 0.8) +
  
  scale_x_continuous(
    "Year",
    breaks = seq(min(yearly_summary$year),
                 max(yearly_summary$year),
                 by = 2),
    expand = expansion(add = c(0.5, 0.5))
  ) +
  
  scale_y_continuous(
    "Proportion Same-Chain Pairs",
    labels = scales::number_format(accuracy = 0.01),
    breaks = scales::pretty_breaks(n = 5),
    expand = expansion(mult = c(0, 0.05))
  ) +
  
  scale_size_area(
    "Cases",
    max_size = 10,
    breaks   = pretty(yearly_summary$cases, n = 4)
  ) +
  
  scale_color_brewer(
    "Serotype",
    palette = "Set1"
  ) +
  guides(
    color = guide_legend(order = 1),
    size  = guide_legend(order = 2, title.position = "top")
  ) +
  
  theme_classic(base_size = 14) +
  theme(
    panel.grid        = element_blank(),
    axis.text         = element_text(color = "black", size = 10),
    axis.title        = element_text(face = "bold"),
    legend.position   = "bottom",
    legend.box        = "vertical",
    legend.title.align= 0.5,
    plot.title        = element_text(face = "bold", hjust = 0.5),
    plot.margin       = margin(5, 5, 5, 5)
  ) +
  
  labs(
    title    = "Same-Chain Proportion vs. Epidemic Size by Serotype",
  )
```


###TEST MATHEMATICALLY 

#Makes a dataframe 
```{r}
glm_df <- yearly_summary %>%
  mutate(
    successes = round(prop_same_chain * n_pairs),  
    failures  = n_pairs - successes,
    log_cases = log(cases)                        
  )
```

#Run Spearman's Test 
```{r}
spearman_results <- glm_df %>%
  group_by(serotype) %>%
  filter(n() >= 3) %>%
  summarize(
    rho = cor(log_cases, prop_same_chain, method = "spearman"),
    p   = cor.test(log_cases, prop_same_chain, method = "spearman")$p.value,
    .groups="drop"
  )

print(spearman_results)
```

#Run a Quasi-Binomial 
```{r}
glm_results <- glm_df %>%
  group_by(serotype) %>%
  filter(n() >= 3) %>% 
  do({
    fit <- glm(
      cbind(successes, failures) ~ log_cases,
      data   = .,
      family = quasibinomial()
    )
    tidy(fit, conf.int = TRUE)
  })

print(glm_results)

glm_results_subset <- glm_results %>%
  filter(term == "log_cases") %>%
  dplyr::select(
    serotype, estimate, std.error, statistic, p.value, conf.low, conf.high
  )

print(glm_results_subset)
```


#Plot GLM Fit 
```{r}
glm_preds <- glm_df %>%
  group_by(serotype) %>%
  filter(n() >= 3) %>%
  nest() %>%
  mutate(
    fit = purrr::map(data, ~ glm(cbind(successes, failures) ~ log_cases, data = ., family = quasibinomial())),
    preds = purrr::map2(fit, data, ~ broom::augment(.x, newdata = .y, type.predict = "response"))
  ) %>%
  unnest(preds) %>%
  ungroup()


ggplot(glm_preds, aes(x = log_cases, y = prop_same_chain)) +
  geom_point(aes(size = n_pairs), alpha = 0.6) +           
  geom_line(aes(y = .fitted), color = "red", size = 1.2) + 
  facet_wrap(~serotype) +                                 
  labs(
    x = "log(Number of Cases)",
    y = "Proportion Same Chain",
    title = "Observed and Fitted Proportion Same Chain vs Epidemic Size by Serotype"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

```{r}
ggplot(yearly_summary, aes(x = year, y = cases, fill = serotype)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~serotype, scales = "free_y") +
  labs(
    title = "Cases Per Year by Serotype",
    x = "Year",
    y = "Number of Cases"
  ) +
  theme_minimal()
```
#Combined All Serotype Data 

```{r}
ggplot(yearly_summary,
       aes(x = year,
           y = prop_same_chain,
           size = cases)) +  # no color or group

  geom_line(size = 0.8, color = "darkblue") +
  geom_point(shape = 16, alpha = 0.8, color = "darkblue") +

  scale_x_continuous(
    "Year",
    breaks = seq(min(yearly_summary$year),
                 max(yearly_summary$year),
                 by = 2),
    expand = expansion(add = c(0.5, 0.5))
  ) +

  scale_y_continuous(
    "Proportion Same-Chain Pairs",
    labels = scales::number_format(accuracy = 0.01),
    breaks = scales::pretty_breaks(n = 5),
    expand = expansion(mult = c(0, 0.05))
  ) +

  scale_size_area(
    "Cases",
    max_size = 10,
    breaks = pretty(yearly_summary$cases, n = 4)
  ) +

  # Remove color scale, keep only size
  guides(
    size = guide_legend(order = 1, title.position = "top")
  ) +

  theme_classic(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.title.align = 0.5,
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.margin = margin(5, 5, 5, 5)
  ) +

  labs(
    title = "Same-Chain Proportion by Epidemic Size",
  )

```

#All serotype Data Mapped
```{r}
ggplot(yearly_summary,
       aes(x = year,
           y = prop_same_chain,
           size = cases)) + 
  geom_point(shape = 16, alpha = 0.8, color = "darkblue") +

  scale_x_continuous(
    "Year",
    breaks = seq(min(yearly_summary$year),
                 max(yearly_summary$year),
                 by = 2),
    expand = expansion(add = c(0.5, 0.5))
  ) +

  scale_y_continuous(
    "Proportion Same-Chain Pairs",
    labels = scales::number_format(accuracy = 0.01),
    breaks = scales::pretty_breaks(n = 5),
    expand = expansion(mult = c(0, 0.05))
  ) +

  scale_size_area(
    "Cases",
    max_size = 10,
    breaks = pretty(yearly_summary$cases, n = 4)
  ) +

  # Remove color scale, keep only size
  guides(
    size = guide_legend(order = 1, title.position = "top")
  ) +

  theme_classic(base_size = 14) +
  theme(
    panel.grid = element_blank(),
    axis.text = element_text(color = "black", size = 10),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom",
    legend.box = "vertical",
    legend.title.align = 0.5,
    plot.title = element_text(face = "bold", hjust = 0.5),
    plot.margin = margin(5, 5, 5, 5)
  ) +

  labs(
    title = "Same-Chain Proportion by Epidemic Size",
  )
```

#Under 5 cases excluded 
```{r}
yearly_summary %>%
  filter(cases >= 5) %>%
  ggplot(aes(x = year,
             y = prop_same_chain,
             size = cases)) + 
    geom_point(shape = 16, alpha = 0.8, color = "darkblue") +
  
    scale_x_continuous(
      "Year",
      breaks = seq(min(yearly_summary$year),
                   max(yearly_summary$year),
                   by = 2),
      expand = expansion(add = c(0.5, 0.5))
    ) +
  
    scale_y_continuous(
      "Proportion Same-Chain Pairs",
      labels = scales::number_format(accuracy = 0.01),
      breaks = scales::pretty_breaks(n = 5),
      expand = expansion(mult = c(0, 0.05))
    ) +
  
    scale_size_area(
      "Cases",
      max_size = 10,
      breaks = pretty(yearly_summary$cases, n = 4)
    ) +
  
    # Keep only size legend
    guides(
      size = guide_legend(order = 1, title.position = "top")
    ) +
  
    theme_classic(base_size = 14) +
    theme(
      panel.grid   = element_blank(),
      axis.text    = element_text(color = "black", size = 10),
      axis.title   = element_text(face = "bold"),
      legend.position    = "bottom",
      legend.box         = "vertical",
      legend.title.align = 0.5,
      plot.title         = element_text(face = "bold", hjust = 0.5),
      plot.margin        = margin(5, 5, 5, 5)
    ) +
  
    labs(
      title = "Same-Chain Proportion by Epidemic Size"
    )

```

```{r}
yearly_summary %>%
  filter(cases >= 5) %>%
  ggplot(aes(x = year,
             y = prop_same_chain,
             size = cases)) + 
    geom_point(shape = 16, alpha = 0.8, color = "darkred") +
  
    scale_x_continuous(
      name   = "Year",
      breaks = seq(min(yearly_summary$year),
                   max(yearly_summary$year),
                   by = 2),
      expand = expansion(add = c(0.5, 0.5))
    ) +
  
    scale_y_continuous(
      name   = "Proportion Same-Chain Pairs",
      labels = scales::number_format(accuracy = 0.01),
      breaks = scales::pretty_breaks(n = 5),
      expand = expansion(mult = c(0, 0.05))
    ) +
  
    scale_size_area(
      name      = "Cases",
      max_size  = 10,
      breaks    = pretty(yearly_summary$cases, n = 4)
    ) +
  
    guides(
      size = guide_legend(
        title.position = "top",
        title.hjust    = 0.5,
        label.position = "right",
        keywidth       = unit(1, "lines"),
        keyheight      = unit(1, "lines"),
        reverse        = FALSE
      )
    ) +
  
    theme_classic(base_size = 14) +
    theme(
      panel.grid         = element_blank(),
      axis.line          = element_line(color = "black"),
      axis.ticks         = element_line(color = "black"),
      axis.text          = element_text(color = "black", size = 10),
      axis.title         = element_text(face = "plain", color = "black"),
      legend.position    = "right",
      legend.direction   = "vertical",
      legend.key.size    = unit(1, "lines"),
      legend.title       = element_text(size = 11, face = "plain"),
      legend.text        = element_text(size = 9),
      legend.background  = element_rect(fill = "white", color = "white"),
      legend.spacing.y   = unit(0.3, "lines"),
      legend.margin      = margin(4, 4, 4, 4)
    )

```




#Run Spearman on All Data 
```{r}
rho_all <- cor(glm_df$log_cases, glm_df$prop_same_chain, method = "spearman")
spearman_test <- cor.test(glm_df$log_cases, glm_df$prop_same_chain, method = "spearman")

cat("Spearman's rho:", rho_all, "\n")
cat("p-value:", spearman_test$p.value, "\n")

```

#Check Normal GLM 
```{r}
fit_binom <- glm(cbind(successes, failures) ~ log_cases,
                 family = binomial(),
                 data = glm_df)
```

#Check Deviance
```{r}
phi_hat2 <- deviance(fit_binom) / fit_binom$df.residual
print(phi_hat2)
```

#Quasi-Binomial GLM 
```{r}
fit_all <- glm(
  cbind(successes, failures) ~ log_cases,
  data   = glm_df,
  family = quasibinomial()
)

glm_results_all <- tidy(fit_all, conf.int = TRUE)

glm_results_or <- glm_results_all %>%
  mutate(
    OR        = exp(estimate),
    OR_CI_low = exp(conf.low),
    OR_CI_high= exp(conf.high)
  ) %>%
  select(
    term, 
    OR, 
    OR_CI_low, 
    OR_CI_high, 
    p.value
  )

print(glm_results_or)
```

#NEW OLRE
```{r}
library(lme4)

glm_df <- glm_df %>% mutate(obs_id = row_number())

glmm_olre <- glmer(
  cbind(successes, failures) ~ log_cases 
    + (1 | serotype) 
    + (1 | obs_id),
  family  = binomial(link = "logit"),
  data    = glm_df,
  control = glmerControl(optimizer = "bobyqa")
)

pr  <- residuals(glmm_olre, type = "pearson")
phi <- sum(pr^2) / df.residual(glmm_olre)
cat("OLRE model dispersion φ̂ =", round(phi, 3), "\n")


```

#Model OUTPUT 
```{r}
summary(glmm_olre)
```

```{r}
preds <- ggpredict(
  model = glmm_olre,
  terms = c("log_cases [all]", "serotype"),
  type  = "random"
)

glm_df <- glm_df %>%
  mutate(p_obs = successes / (successes + failures))

ggplot() +
  geom_jitter(
    data = glm_df,
    aes(x = log_cases, y = p_obs, colour = serotype),
    width  = 0,
    height = 0.02,
    alpha  = 0.6,
    size   = 1.5
  ) +
  geom_line(
    data = preds,
    aes(x = x, y = predicted, colour = group),
    size = 1
  ) +
  geom_ribbon(
    data = preds,
    aes(x = x, ymin = conf.low, ymax = conf.high, fill = group),
    alpha = 0.2, colour = NA
  ) +
  scale_colour_brewer("Serotype", palette = "Set1") +
  scale_fill_brewer("Serotype", palette = "Set1") +
  labs(
    x        = "log(cases)",
    y        = "Proportion of successes",
    title    = "GLMM Fit with Outbreak-Level Data Points",
    subtitle = "Points = observed proportions; Curves = fitted serotype-specific predictions"
  ) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "right")
```

```{r}
glm_df <- glm_df %>%
  mutate(
    p_obs = successes / (successes + failures),
    N     = successes + failures
  )

preds_fixed <- ggpredict(
  model = glmm_olre,
  terms = "log_cases [all]",
  type  = "fixed"
)

ggplot() +
  geom_point(
    data     = glm_df,
    aes(x = log_cases, y = p_obs, size = N),
    shape    = 21,
    colour   = "red",
    fill     = "red",
    alpha    = 0.6,
    position = position_jitter(width = 0, height = 0.02)
  ) +
  scale_size_area(
    name     = "N cases",
    max_size = 6
  ) +
  geom_line(
    data   = preds_fixed,
    aes(x = x, y = predicted),
    colour = "steelblue",
    size   = 1.2
  ) +
  geom_ribbon(
    data   = preds_fixed,
    aes(x = x, ymin = conf.low, ymax = conf.high),
    fill   = "steelblue",
    alpha  = 0.2,
    inherit.aes = FALSE
  ) +
  scale_x_continuous("log(cases)", expand = expansion(add = 0.1)) +
  scale_y_continuous("Predicted P(success)",
                     limits = c(0,1),
                     expand = expansion(add = 0.02)) +
  theme_classic(base_size = 14) +
  theme(
    panel.grid     = element_blank(),
    axis.line      = element_line(colour = "black"),
    axis.ticks     = element_line(colour = "black"),
    axis.title     = element_text(face = "plain"),
    axis.text      = element_text(face = "plain"),
    legend.position = "right"
  )

```


```{r}
glm_df <- glm_df %>%
  mutate(
    p_obs = successes / (successes + failures),
    N     = successes + failures
  )

preds_fixed <- ggpredict(
  model = glmm_olre,
  terms = "log_cases [all]",
  type  = "fixed"
)

preds_random <- ggpredict(
  model = glmm_olre,
  terms = c("log_cases [all]", "serotype"),
  type  = "random"
)

ggplot() +
  geom_line(
    data   = preds_random,
    aes(x = x, y = predicted, group = group),
    colour = "#4A90E2",   # darker blue
    size   = 0.8,         # a bit thicker
    alpha  = 0.7          # more opaque
  ) +
  geom_point(
    data     = glm_df,
    aes(x = log_cases, y = p_obs, size = N),
    shape    = 21,
    fill     = "red",
    colour   = "red",
    stroke   = 0.3,
    alpha    = 0.7
  ) +
  scale_size_area(name = "N cases", max_size = 6) +
  geom_line(
    data   = preds_fixed,
    aes(x = x, y = predicted),
    colour = "#2C3E50",
    size   = 1.2
  ) +
  geom_ribbon(
    data   = preds_fixed,
    aes(x = x, ymin = conf.low, ymax = conf.high),
    fill   = "#2C3E50",
    alpha  = 0.2,
    inherit.aes = FALSE
  ) +
  scale_x_continuous("Logarithm of case count", expand = expansion(add = 0.1)) +
  scale_y_continuous("Predicted Probability of Success",
                     limits = c(0,1),
                     expand = expansion(add = 0.02)) +
  theme_classic(base_size = 14) +
  theme(
    panel.grid      = element_blank(),
    axis.line       = element_line(colour = "black"),
    axis.ticks      = element_line(colour = "black"),
    axis.title      = element_text(face = "plain"),
    axis.text       = element_text(face = "plain"),
    legend.position = "right"
  )

```

```{r}
ggplot() +
  geom_line(
    data   = preds_random,
    aes(x = x, y = predicted, group = group),
    colour = "#4A90E2",
    size   = 0.8,
    alpha  = 0.7
  ) +
  geom_point(
    data     = glm_df,
    aes(x = log_cases, y = p_obs, size = N),
    shape    = 21,
    fill     = "darkred",
    colour   = "darkred",
    stroke   = 0.3,
    alpha    = 0.7
  ) +
  scale_size_area(max_size = 6) +
  geom_line(
    data   = preds_fixed,
    aes(x = x, y = predicted),
    colour = "#2C3E50",
    size   = 1.2
  ) +
  geom_ribbon(
    data   = preds_fixed,
    aes(x = x, ymin = conf.low, ymax = conf.high),
    fill   = "#2C3E50",
    alpha  = 0.2,
    inherit.aes = FALSE
  ) +
  scale_x_continuous("Logarithm of case count", expand = expansion(add = 0.1)) +
  scale_y_continuous("Predicted probability of success",
                     limits = c(0,1),
                     expand = expansion(add = 0.02)) +
  theme_classic(base_size = 14) +
    theme(
    panel.grid      = element_blank(),
    axis.line       = element_line(colour = "black"),
    axis.ticks      = element_line(colour = "black"),
    axis.title      = element_text(face = "plain",  color = "black"),
    axis.text       = element_text(face = "plain",  color = "black"),
    strip.text      = element_text(color = "black"),     # facet labels
    plot.title      = element_text(color = "black"),     # if you ever add one
    legend.text     = element_text(color = "black"),
    legend.title    = element_text(color = "black"),
    legend.position = "none"
  )
```

```{r}
ggplot() +
  geom_line(
    data   = preds_random,
    aes(x = x, y = predicted, group = group),
    colour = "#4A90E2",
    size   = 0.8,
    alpha  = 0.7
  ) +
  # raw data points
  geom_point(
    data     = glm_df,
    aes(x = log_cases, y = p_obs, size = N),
    shape    = 21,
    fill     = "darkred",
    colour   = "darkred",
    stroke   = 0.3,
    alpha    = 0.7
  ) +
  scale_size_area(max_size = 6) +
  # fixed‐effects curve + CI
  geom_line(
    data   = preds_fixed,
    aes(x = x, y = predicted),
    colour = "#2C3E50",
    size   = 1.2
  ) +
  geom_ribbon(
    data   = preds_fixed,
    aes(x = x, ymin = conf.low, ymax = conf.high),
    fill   = "#2C3E50",
    alpha  = 0.2,
    inherit.aes = FALSE
  ) +
  # axes labels
  scale_x_continuous(
    "Logarithm of case count",
    expand = expansion(add = 0.1)
  ) +
  scale_y_continuous(
    "Probability of same chain",
    limits = c(0, 0.5),
    expand = expansion(add = 0.02)
  ) +
  # clean theme, no grid, no legend
  theme_classic(base_size = 14) +
  theme(
    panel.grid      = element_blank(),
    axis.line       = element_line(colour = "black"),
    axis.ticks      = element_line(colour = "black"),
    axis.title      = element_text(face = "plain",  color = "black"),
    axis.text       = element_text(face = "plain",  color = "black"),
    strip.text      = element_text(color = "black"),     # facet labels
    plot.title      = element_text(color = "black"),     # if you ever add one
    legend.text     = element_text(color = "black"),
    legend.title    = element_text(color = "black"),
    legend.position = "none"
  )

```

```{r}
library(lme4)

obs_or <- exp(fixef(glmm_olre)["log_cases"])

set.seed(123)  
boot_or <- bootMer(
  glmm_olre,
  FUN = function(fit) exp(fixef(fit)["log_cases"]),
  nsim = 1000,
  type = "parametric"
)

ci <- quantile(boot_or$t, probs = c(0.025, 0.975))

cat("Observed OR:", round(obs_or, 3), "\n")
cat("95% CI: [", round(ci[1], 3), ", ", round(ci[2], 3), "]\n")
```


```{r}
boot_df <- glm_df  

boot_stat <- function(data, indices) {
  d <- data[indices, ]
  fit <- glmer(
    cbind(successes, failures) ~ log_cases + (1 | serotype) + (1 | obs_id),
    family  = binomial(link = "logit"),
    data    = d,
    control = glmerControl(optimizer = "bobyqa")
  )
  return(exp(fixef(fit)["log_cases"]))
}

set.seed(123)
boot_res_np <- boot(
  data  = boot_df,
  statistic = boot_stat,
  R     = 1000
)

ci_np <- boot.ci(boot_res_np, type = "perc")$percent[4:5]

obs_or <- exp(fixef(glmm_olre)["log_cases"])
cat("Observed OR:", round(obs_or, 3), "\n")
cat("95% CI (non-parametric bootstrap): [", 
    round(ci_np[1], 3), ", ", round(ci_np[2], 3), "]\n")

```

```{r}
beta_hat   <- -0.5887             
ci_log     <- c(-0.849, -0.324)    

# Exponentiate to get odds ratio and its CI
or_hat     <- exp(beta_hat)
ci_or      <- exp(ci_log)

cat(sprintf("OR = %.3f\n95%% CI = [%.3f, %.3f]\n",
            or_hat, ci_or[1], ci_or[2]))

```

```{r}
beta0 <- fixef(glmm_olre)["(Intercept)"]

ci_log0 <- confint(glmm_olre, parm="(Intercept)", method="Wald")

or0    <- exp(beta0)
ci_or0 <- exp(ci_log0)

cat(sprintf(
  "Intercept OR = %.3f\n95%% CI = [%.3f, %.3f]\n",
  or0, ci_or0[1], ci_or0[2]
))

```


#SOME DIAGNOSTIC METRICS 
```{r}
library(lme4)
library(DHARMa)
library(performance)
library(lmtest)
library(dplyr)

glm_df <- glm_df %>% mutate(obs_id = row_number())

glmm_olre <- glmer(
  cbind(successes, failures) ~ log_cases + (1 | serotype) + (1 | obs_id),
  family  = binomial(link = "logit"),
  data    = glm_df,
  control = glmerControl(optimizer = "bobyqa")
)

summary(glmm_olre)

fit_obs_only <- glmer(
  cbind(successes, failures) ~ log_cases + (1 | obs_id),
  family = binomial(link = "logit"),
  data   = glm_df,
  control= glmerControl(optimizer="bobyqa")
)
anova(fit_obs_only, glmm_olre, test="Chisq")

fit_sero_only <- glmer(
  cbind(successes, failures) ~ log_cases + (1 | serotype),
  family = binomial(link = "logit"),
  data   = glm_df,
  control= glmerControl(optimizer="bobyqa")
)
anova(fit_sero_only, glmm_olre, test="Chisq")

pr <- residuals(glmm_olre, type="pearson")
phi_hat <- sum(pr^2) / df.residual(glmm_olre)
cat("Dispersion φ̂ =", round(phi_hat, 3), "\n")


print(r2_nakagawa(glmm_olre))


sim_res <- simulateResiduals(glmm_olre, n = 1000)
plot(sim_res)                     # overall QQ & residual vs predicted
plotResiduals(sim_res, glm_df$log_cases)


dotchart(ranef(glmm_olre)$serotype[,1], main="Serotype RE intercepts")
dotchart(ranef(glmm_olre)$obs_id[,1],    main="Obs‐ID RE intercepts")

```

#LRT 
```{r}
library(lmtest)

fit_obs_only <- glmer(
  cbind(successes, failures) ~ log_cases + (1 | obs_id),
  family = binomial(link="logit"),
  data   = glm_df,
  control= glmerControl(optimizer="bobyqa")
)

fit_sero_only <- glmer(
  cbind(successes, failures) ~ log_cases + (1 | serotype),
  family = binomial(link="logit"),
  data   = glm_df,
  control= glmerControl(optimizer="bobyqa")
)

glmm_olre <- glmer(
  cbind(successes, failures) ~ log_cases + (1 | serotype) + (1 | obs_id),
  family = binomial(link="logit"),
  data   = glm_df,
  control= glmerControl(optimizer="bobyqa")
)

lrtest(fit_obs_only, glmm_olre)
lrtest(fit_sero_only, glmm_olre)
```


#Graph Prep 
```{r}
glm_preds_all <- broom::augment(fit_all, newdata = glm_df, type.predict = "response")

glm_preds_all <- glm_df %>% bind_cols(.fitted = glm_preds_all$.fitted)

```

#GRAPH OF MODEL 
```{r}
ggplot(glm_preds_all, aes(x = log_cases, y = prop_same_chain)) +
  geom_point(aes(size = cases), color = "#08519c", alpha = 0.55) +
  geom_line(aes(y = .fitted), color = "#d73027", size = 1.2) +
  scale_size_continuous(
    name = "Cases",
    range = c(2.5, 7),
    breaks = pretty(glm_preds_all$cases, n = 3)
  ) +
  scale_x_continuous(
    breaks = scales::pretty_breaks(n = 5)
  ) +
  scale_y_continuous(
    limits = c(0, 0.5),
    breaks = seq(0, 0.5, by = 0.1),
    expand = expansion(mult = c(0, 0.03))
  ) +
  labs(
    x = "log(Number of Cases)",
    y = "Proportion Same-Chain Pairs"
  ) +
  theme_minimal(base_size = 15) +
  theme(
    axis.title      = element_text(face = "bold", size = 15),
    axis.text       = element_text(size = 13, color = "black"),
    plot.title      = element_text(face = "bold", size = 16, hjust = 0.5),
    legend.position = "bottom",
    legend.title    = element_text(size = 12),
    legend.text     = element_text(size = 11),
    # Remove gridlines
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    # Add black axis lines and ticks for professional look
    axis.line        = element_line(size = 1, color = "black"),
    axis.ticks       = element_line(size = 0.8, color = "black"),
    plot.margin      = margin(10, 10, 10, 10)
  )
```

