---
title: "figure_generation"
author:
date: "`r Sys.Date()`"
output: html_document
---
#Load libraries
```{r}
library(librarian)
shelf(dplyr)
shelf(tidyr)
shelf(tibble)
shelf(ggplot2)
shelf(stringr)
shelf(scales)
shelf(patchwork)
shelf(binom)
shelf(ape)
shelf
```

#Find common samples between all three matrices created 
```{r}
#DENV4 COMMON LENGTHS 
#105 out of 119
d4_patristic_cophen_id  <- rownames(d4_patristic_cophen)
d4_dist_df_id <- rownames(d4_distance)
d4_time_df_id <-rownames(d4_time)
common_ids_4 <- Reduce(intersect, list(d4_patristic_cophen_id, d4_dist_df_id, d4_time_df_id))
length(common_ids_4)

#DENV3 COMMON LENGTHS 
#140 out of 159
d3_patristic_cophen_id  <- rownames(d3_patristic_cophen)
d3_dist_df_id <- rownames(d3_distance)
d3_time_df_id <-rownames(d3_time)
common_ids_3 <- Reduce(intersect, list(d3_patristic_cophen_id, d3_dist_df_id, d3_time_df_id))
length(common_ids_3)

#DENV1 COMMON LENGTHS 
#264 out of 291
d1_patristic_cophen_id  <- rownames(d1_patristic_cophen)
d1_dist_df_id <- rownames(d1_distance)
d1_time_df_id <-rownames(d1_time)
common_ids_1 <- Reduce(intersect, list(d1_patristic_cophen_id, d1_dist_df_id, d1_time_df_id))
length(common_ids_1)

#DENV2 COMMON LENGTHS 
#267 out of 282
d2_patristic_cophen_id  <- rownames(d2_patristic_cophen)
d2_dist_df_id <- rownames(d2_distance)
d2_time_df_id <-rownames(d2_time)
common_ids_2 <- Reduce(intersect, list(d2_patristic_cophen_id, d2_dist_df_id, d2_time_df_id))
length(common_ids_2)
```

#Subset Matrices
```{r}
#Subset the matrices now D4
d4_patristic_sub <- d4_patristic_cophen[common_ids_4, common_ids_4]
d4_dist_sub      <- d4_distance[common_ids_4, common_ids_4]
d4_time_sub      <- d4_time[common_ids_4, common_ids_4]

#Subset the matrices now D3
d3_patristic_sub <- d3_patristic_cophen[common_ids_3, common_ids_3]
d3_dist_sub      <- d3_distance[common_ids_3, common_ids_3]
d3_time_sub      <- d3_time[common_ids_3, common_ids_3]

#Subset the matrices now D1
d1_patristic_sub <- d1_patristic_cophen[common_ids_1, common_ids_1]
d1_dist_sub      <- d1_distance[common_ids_1, common_ids_1]
d1_time_sub      <- d1_time[common_ids_1, common_ids_1]

#Subset the matrices now D2
d2_patristic_sub <- d2_patristic_cophen[common_ids_2, common_ids_2]
d2_dist_sub      <- d2_distance[common_ids_2, common_ids_2]
d2_time_sub      <- d2_time[common_ids_2, common_ids_2]
```

#Source helper merging data frame 
```{r}
source("script/clean/helper_scripts/helper_merging_df.R")
```

#Pairs DF Formation
```{r}
d4_pairs_df <- fc_combined_df(d4_patristic_sub,d4_dist_sub,d4_time_sub)
d3_pairs_df <- fc_combined_df(d3_patristic_sub,d3_dist_sub,d3_time_sub)
d1_pairs_df <- fc_combined_df(d1_patristic_sub,d1_dist_sub,d1_time_sub)
d2_pairs_df <- fc_combined_df(d2_patristic_sub,d2_dist_sub,d2_time_sub)
```

#Add a generation time axis 
```{r}
generation_time_years <- 20 / 365  #0.05479 years per dengue generation for 20 days 
```


#Plot for Figure 1G: Spatial Distance (km) vs Evolutionary Time (years) for DENV1 serotype 

#Overall Median Line for all plots 
```{r}
fc_median_pd <- function(df) {
  df %>%
    pull(spatial_dist) %>%
    median(na.rm = TRUE)
}

d1_overall_median <- fc_median_pd(d1_pairs_df)
d2_overall_median <- fc_median_pd(d2_pairs_df)
d3_overall_median <- fc_median_pd(d3_pairs_df)
d4_overall_median <- fc_median_pd(d4_pairs_df)

```

#Source helper function
```{r}
source("script/clean/helper_scripts/helper_1G_binning.R")
```

#D1
```{r}
d1_binned_df <- fc_1G_binning(d1_pairs_df) 

d1_evol_time <- ggplot(d1_binned_df, aes(x = evolutionary_time, y = spatial_mean)) +
  geom_ribbon(aes(ymin = spatial_lower, ymax = spatial_upper), fill = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = d1_overall_median, linetype = "dashed", color = "black") +
  
  scale_x_continuous(
    name = "Evolutionary time (years)",
    sec.axis = sec_axis(
      trans = ~ . / generation_time_years, 
      name = "Transmission generations (No.)"
    )
  ) +
  theme_minimal() +
  theme(
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top = element_text(size = 10),
    axis.line.x.top = element_line(color = "black")
  ) +
  
  labs(
    y = "Spatial distance (km)"
  )

print(d1_evol_time)
#ggsave("d1_1G.pdf", plot = d1_evol_time, device = "pdf", width = 8, height = 6)
```

#D4
```{r}
d4_binned_df <-  fc_1G_binning(d4_pairs_df) 

d4_evol_time <- ggplot(d4_binned_df, aes(x = evolutionary_time, y = spatial_mean)) +
  geom_ribbon(aes(ymin = spatial_lower, ymax = spatial_upper), fill = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = d4_overall_median, linetype = "dashed", color = "black") +

  scale_x_continuous(
    name = "Evolutionary time (years)",
    sec.axis = sec_axis(
      trans = ~ . / generation_time_years,
      name = "Transmission generations (No.)"
    )
  ) +

  theme_minimal() +
  theme(
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top = element_text(size = 10),
    axis.line.x.top = element_line(color = "black")
  ) +

  labs(
    y = "Spatial distance (km)"
  )

print(d4_evol_time)

#ggsave("d4_1G.pdf", plot = d4_evol_time, device = "pdf", width = 8, height = 6)
```

#D3 
```{r}
d3_binned_df <- fc_1G_binning(d3_pairs_df) 

d3_evol_time <- ggplot(d3_binned_df, aes(x = evolutionary_time, y = spatial_mean)) +
  geom_ribbon(aes(ymin = spatial_lower, ymax = spatial_upper), fill = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = d3_overall_median, linetype = "dashed", color = "black") +
  
  scale_x_continuous(
    name = "Evolutionary time (years)",
    sec.axis = sec_axis(
      trans = ~ . / generation_time_years, 
      name = "Transmission generations (No.)"
    )
  ) +
  theme_minimal() +
  theme(
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top = element_text(size = 10),
    axis.line.x.top = element_line(color = "black")
  ) +
  
  labs(
    y = "Spatial distance (km)"
  )

print(d3_evol_time)
#ggsave("d3_1G.pdf", plot = d3_evol_time, device = "pdf", width = 8, height = 6)
```

#D2 
```{r}
d2_binned_df <- fc_1G_binning(d2_pairs_df) 

d2_evol_time <- ggplot(d2_binned_df, aes(x = evolutionary_time, y = spatial_mean)) +
  geom_ribbon(aes(ymin = spatial_lower, ymax = spatial_upper), fill = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = d2_overall_median, linetype = "dashed", color = "black") +
  scale_x_continuous(
    name = "Evolutionary time (years)",
    sec.axis = sec_axis(
      trans = ~ . / generation_time_years, 
      name = "Transmission generations (No.)"
    )
  ) +
  theme_minimal() +
  theme(
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top = element_text(size = 10),
    axis.line.x.top = element_line(color = "black")
  ) +
  labs(
    y = "Spatial distance (km)"
  )
print(d2_evol_time)
#ggsave("d2_1G.pdf", plot = d2_evol_time, device = "pdf", width = 8, height = 6)
```

#Combined Figure 1G
```{r}
#edit each binned df with a name 
d3_binned_df <- d3_binned_df %>%
  mutate(serotype = "DENV-3")

d4_binned_df <- d4_binned_df %>%
  mutate(serotype = "DENV-4")

d1_binned_df <- d1_binned_df %>%
  mutate(serotype = "DENV-1")

d2_binned_df <- d2_binned_df %>%
  mutate(serotype = "DENV-2")

#make a total binned data frame 
combined_binned_df <- bind_rows(d3_binned_df, d4_binned_df, d1_binned_df, d2_binned_df)

#make a total median 
combined_overall_median <- bind_rows(
  d1_pairs_df,
  d2_pairs_df,
  d3_pairs_df,
  d4_pairs_df
) %>%
  pull(spatial_dist) %>%
  median(na.rm = TRUE)


combined_evol_time <- ggplot(combined_binned_df, aes(x = evolutionary_time, y = spatial_mean, color = serotype, fill = serotype)) +
  geom_ribbon(aes(ymin = spatial_lower, ymax = spatial_upper), alpha = 0.3, color = NA) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  geom_hline(yintercept = combined_overall_median, linetype = "dashed", color = "black") +
  
  # Updated color and fill manual scales to include DENV-2 (set to purple here)
  scale_color_manual(values = c("DENV-3" = "blue", "DENV-4" = "darkred", "DENV-1" = "darkgreen", "DENV-2" = "purple")) +
  scale_fill_manual(values = c("DENV-3" = "blue", "DENV-4" = "darkred", "DENV-1" = "darkgreen", "DENV-2" = "purple")) +
  
  scale_x_continuous(
    name = "Evolutionary time (years)",
    sec.axis = sec_axis(
      trans = ~ . / generation_time_years,
      name = "Transmission generations (No.)"
    )
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top = element_text(size = 10),
    axis.line.x.top = element_line(color = "black"),
    legend.title = element_blank(),
    legend.position = "top"
  ) +
  
  labs(
    y = "Spatial distance (km)"
  )

print(combined_evol_time)

#ggsave("combined_1G.pdf", plot = combined_evol_time, device = "pdf", width = 8, height = 6)
```

##Plot for Figure 2A: Proportion Same Chain vs Distance (km) for DENV1 serotype 

#Source Same Chain
```{r}
source("script/clean/helper_scripts/helper_same_chain.R")
```

#Function for same chain 
```{r}
d1_pairs_df <- fc_same_chain(d1_pairs_df)
d2_pairs_df <- fc_same_chain(d2_pairs_df)
d3_pairs_df <- fc_same_chain(d3_pairs_df)
d4_pairs_df <- fc_same_chain(d4_pairs_df)
```

#Total Same Chains
```{r}
sum(d1_pairs_df$same_chain, na.rm = TRUE)
sum(d2_pairs_df$same_chain, na.rm = TRUE)
sum(d3_pairs_df$same_chain, na.rm = TRUE)
sum(d4_pairs_df$same_chain, na.rm = TRUE)
```
#Source 2A Binning Function
```{r}
source("script/clean/helper_scripts/helper_2A_binning.R")
```

#Make all the Binned Data Frames 
```{r}
d1_binned_chain_df <- fc_2A_binning(d1_pairs_df)
d2_binned_chain_df <- fc_2A_binning(d2_pairs_df)
d3_binned_chain_df <- fc_2A_binning(d3_pairs_df)
d4_binned_chain_df <- fc_2A_binning(d4_pairs_df)
```

#DENV3 Graph Attempt
```{r}
d3_prop_chain <- ggplot(
    d3_binned_chain_df,
    aes(x = bin_cutoff, y = prop_same_chain)
  ) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    width = 0.2, color = "purple"
  ) +
  geom_point(
    shape = 21, size = 3, stroke = 1.2,
    color = "purple", fill = "white"
  ) +
  scale_y_continuous(
    name   = "Proportion Same Chain",
    breaks = seq(0, 1, by = 0.02),
    labels = number_format(accuracy = 0.01),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, by = 5),
    labels = number_format(accuracy = 1)
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor.y = element_line(color = "gray95"),
    panel.grid.x       = element_blank(),
    axis.line          = element_line(size = 0.6, color = "black"),
    axis.ticks         = element_line(size = 0.6, color = "black"),
    axis.title         = element_text(size = 14, face = "bold"),
    axis.text.x        = element_text(angle = 45, hjust = 1),
    axis.text          = element_text(size = 12),

    # Publication‐ready title styling
    plot.title         = element_text(
                           hjust = 0.5,          # center
                           face  = "bold",       # bold
                           size  = 18            # larger size
                         )
  ) +
  labs(
    title = "DENV-3",
    y     = "Proportion Same Chain"
  )

print(d3_prop_chain)
```


#DENV4 Graph Attempt
```{r}
d4_prop_chain <- ggplot(
    d4_binned_chain_df,
    aes(x = bin_cutoff, y = prop_same_chain)
  ) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    width = 0.2, color = "purple"
  ) +
  geom_point(
    shape = 21, size = 3, stroke = 1.2,
    color = "purple", fill = "white"
  ) +
  scale_y_continuous(
    name   = "Proportion Same Chain",
    breaks = seq(0, 1, by = 0.01),
    labels = number_format(accuracy = 0.01),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, by = 5),
    labels = number_format(accuracy = 1)
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor.y = element_line(color = "gray95"),
    panel.grid.x       = element_blank(),
    axis.line          = element_line(size = 0.6, color = "black"),
    axis.ticks         = element_line(size = 0.6, color = "black"),
    axis.title         = element_text(size = 14, face = "bold"),
    axis.text.x        = element_text(angle = 45, hjust = 1),
    axis.text          = element_text(size = 12),
    plot.title         = element_text(
                           hjust = 0.5,          # center
                           face  = "bold",       # bold
                           size  = 18            # larger size
                         )
  ) +
  labs(
    title = "DENV-4",
    y     = "Proportion Same Chain"
  )

print(d4_prop_chain)
#ggsave("d4_2A.pdf", plot = d4_prop_chain, device = "pdf", width = 8, height = 6)
```

#DENV1 Graph Attempt
```{r}
d1_prop_chain <- ggplot(
    d1_binned_chain_df,
    aes(x = bin_cutoff, y = prop_same_chain)
  ) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    width = 0.2, color = "purple"
  ) +
  geom_point(
    shape = 21, size = 3, stroke = 1.2,
    color = "purple", fill = "white"
  ) +
  scale_y_continuous(
    name   = "Proportion Same Chain",
    breaks = seq(0, 1, by = 0.01),
    labels = number_format(accuracy = 0.01),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, by = 5),
    labels = number_format(accuracy = 1)
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor.y = element_line(color = "gray95"),
    panel.grid.x       = element_blank(),
    axis.line          = element_line(size = 0.6, color = "black"),
    axis.ticks         = element_line(size = 0.6, color = "black"),
    axis.title         = element_text(size = 14, face = "bold"),
    axis.text.x        = element_text(angle = 45, hjust = 1),
    axis.text          = element_text(size = 12),
    plot.title         = element_text(
                           hjust = 0.5,          # center
                           face  = "bold",       # bold
                           size  = 18            # larger size
                         )
  ) +
  labs(
    title = "DENV-1",
    y     = "Proportion Same Chain"
  )

print(d1_prop_chain)
```

#DENV2 Graph Attempt
```{r}
d2_prop_chain <- ggplot(
    d2_binned_chain_df,
    aes(x = bin_cutoff, y = prop_same_chain)
  ) +
  geom_errorbar(
    aes(ymin = lower, ymax = upper),
    width = 0.2, color = "purple"
  ) +
  geom_point(
    shape = 21, size = 3, stroke = 1.2,
    color = "purple", fill = "white"
  ) +
  scale_y_continuous(
    name = "Proportion Same Chain",
    limits = c(0, 0.03),
    breaks = seq(0, 0.03, by = 0.005),
    expand = c(0, 0)
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, by = 5),
    labels = number_format(accuracy = 1)
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.y = element_line(color = "gray90"),
    panel.grid.minor.y = element_line(color = "gray95"),
    panel.grid.x       = element_blank(),
    axis.line          = element_line(size = 0.6, color = "black"),
    axis.ticks         = element_line(size = 0.6, color = "black"),
    axis.title         = element_text(size = 14, face = "bold"),
    axis.text.x        = element_text(angle = 45, hjust = 1),
    axis.text          = element_text(size = 12),
    plot.title         = element_text(
                           hjust = 0.5,          # center
                           face  = "bold",       # bold
                           size  = 18            # larger size
                         )
  ) +
  labs(
    title = "DENV-2",
    y     = "Proportion Same Chain"
  )

print(d2_prop_chain)

#ggsave("d2_2A.pdf", plot = d2_prop_chain, device = "pdf", width = 8, height = 6)
```

#Combined figure 2A
```{r}
d3_binned_chain_df <- d3_binned_chain_df %>% mutate(serotype = "DENV-3")
d4_binned_chain_df <- d4_binned_chain_df %>% mutate(serotype = "DENV-4")
d1_binned_chain_df <- d1_binned_chain_df %>% mutate(serotype = "DENV-1")
d2_binned_chain_df <- d2_binned_chain_df %>% mutate(serotype = "DENV-2")

combined_df <- bind_rows(d3_binned_chain_df, d4_binned_chain_df,
                         d1_binned_chain_df, d2_binned_chain_df)

max_dist_combined <- max(combined_df$bin_mid, na.rm = TRUE)
```

```{r}
combined_prop_chain <- ggplot(combined_df, aes(x = bin_cutoff, y = prop_same_chain,
                                               color = serotype, fill = serotype)) +
  geom_point(aes(shape = serotype), size = 3, stroke = 1.2) +
  geom_errorbar(aes(ymin = lower, ymax = upper), width = 0.2, alpha = 0.8) +
  
  scale_y_continuous(
    name   = "Proportion Same Chain",
    breaks = seq(0, 1, by = 0.02),
    labels = scales::number_format(accuracy = 0.01),
    expand = expansion(mult = c(0, 0.05))
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, by = 5),
    labels = scales::number_format(accuracy = 1)
  ) +
  
  scale_color_manual(values = c(
    "DENV-3" = "blue", 
    "DENV-4" = "darkred", 
    "DENV-1" = "darkgreen", 
    "DENV-2" = "purple"
  )) +
  scale_fill_manual(values = c(
    "DENV-3" = "white", 
    "DENV-4" = "white", 
    "DENV-1" = "white", 
    "DENV-2" = "white"
  )) +
  scale_shape_manual(values = c(
    "DENV-3" = 21, 
    "DENV-4" = 22, 
    "DENV-1" = 23, 
    "DENV-2" = 24
  )) +
  
  theme_minimal(base_size = 14) +
  theme(
    panel.grid        = element_blank(),
    axis.line         = element_line(size = 0.6, color = "black"),
    axis.ticks        = element_line(size = 0.6, color = "black"),
    axis.title        = element_text(size = 14, face = "plain", color = "black"),
    axis.text         = element_text(size = 12, face = "plain", color = "black"),
    axis.text.x       = element_text(angle = 45, hjust = 1),
    legend.position       = c(0.95, 0.95),
    legend.justification  = c(1, 1),
    legend.title          = element_blank(),
    legend.text           = element_text(color = "black"),
    legend.background     = element_blank(),
    plot.margin           = margin(8, 8, 8, 8)
  ) +
  labs(
    y     = "Proportion Same Chain"
  )

combined_prop_chain
```



```{r}
library(dplyr)
library(purrr)
library(tidyr)
library(minpack.lm)

fit_decay <- function(df){
  start_vals <- list(
    c = min(df$prop_same_chain, na.rm=TRUE),
    a = max(df$prop_same_chain, na.rm=TRUE) - min(df$prop_same_chain, na.rm=TRUE),
    b = 0.1
  )
  nlsLM(
    prop_same_chain ~ c + a * exp(-b * bin_cutoff),
    data    = df,
    start   = start_vals,
    weights = 1 / (df$se^2)
  )
}

serotype_summary <- combined_df %>%
  group_by(serotype) %>%
  nest() %>%
  mutate(
    model = map(data, fit_decay),
    params = map(model, ~{
      cf <- coef(.x)
      CI <- confint(.x, level=0.95)
      bind_rows(
        tibble(
          parameter = c("c","a","b"),
          estimate = cf[c("c","a","b")],
          lower    = CI[c("c","a","b"), 1],
          upper    = CI[c("c","a","b"), 2]
        ),
        tibble(
          parameter = "p0",
          estimate = cf["c"] + cf["a"],
          lower    = CI["c",1] + CI["a",1],
          upper    = CI["c",2] + CI["a",2]
        ),
        tibble(
          parameter    = "half_distance",
          estimate    = log(2)/cf["b"],
          # invert b‐CI for half‐distance
          lower       = log(2)/CI["b", 2],
          upper       = log(2)/CI["b", 1]
        )
      )
    })
  ) %>%
  select(serotype, params) %>%
  unnest(params)

serotype_summary %>% 
  arrange(serotype, parameter) %>%
  mutate_at(vars(estimate:upper), ~round(., 4)) %>%
  print()


```
#Percent Decrease Calculations EVEN ALL SEROTYPES
```{r}
library(dplyr)

serotype_params <- tibble::tribble(
  ~serotype,       ~b,      ~b_low,   ~b_high,
  "All serotypes", 0.1127,  0.1091,   0.1165,
  "DENV-1",        0.1113,  0.1066,   0.1162,
  "DENV-2",        0.1083,  0.1039,   0.1129,
  "DENV-3",        0.0973,  0.0936,   0.1011,
  "DENV-4",        0.1410,  0.1321,   0.1504
)

serotype_params %>% 
  mutate(
    loss_per_km      = (1 - exp(-b))      * 100,
    loss_per_km_low  = (1 - exp(-b_high)) * 100,
    # CI upper bound (best‐case: use b_low)
    loss_per_km_high = (1 - exp(-b_low))  * 100
  ) %>% 
  select(serotype, loss_per_km, loss_per_km_low, loss_per_km_high) %>% 
  mutate(across(-serotype, ~round(.x, 2)))

```

