---
title: "figure_1G_final"
author: 
date: "`r Sys.Date()`"
output: html_document
---
#Plot for 1G FIXED 
```{r}
source("script/clean/helper_scripts/helper_1G_fixed_median.R")
```

#Load Medians 
```{r}
overall_median <- all_data_df %>% 
  summarise(median_spatial = median(spatial_dist, na.rm = TRUE)) %>% 
  pull(median_spatial)

overall_median
```
#All serotypes binning 
```{r}
all_serotypes_binning_1G_fixed <- fc_1G_binning_fixed_median(all_data_df) 
```

#Graph of serotypes 
```{r}
all_serotype_evol_time_fixed <- ggplot(all_serotypes_binning_1G_fixed, aes(x = midpoint, y = spatial_median)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
              fill  = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = overall_median,
             linetype = "dashed",
             color    = "black") +
  scale_x_continuous(
    name    = "Evolutionary time (years)",
    limits  = c(0, 15),
    breaks  = seq(0, 15, by = 1),
    sec.axis = sec_axis( 
      trans = ~ . / generation_time_years,
      name  = "Transmission generations (No.)"
    )
  ) +
  theme_minimal() +
  theme(
    # center the title and make it bold
    plot.title = element_text(hjust = 0.5, face = "bold"),
    # keep your existing x-axis–top styling
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top  = element_text(size = 10),
    axis.line.x.top  = element_line(color = "black")
  ) +
  labs(
    y = "Spatial distance (km)",
    title = "All Serotypes"
  )

all_serotype_evol_time_fixed
```

```{r}
library(ggplot2)

all_serotype_evol_time_fixed <- ggplot(all_serotypes_binning_1G_fixed, 
                                       aes(x = midpoint, y = spatial_median)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
              fill  = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = overall_median,
             linetype = "dashed", color = "black") +
  scale_x_continuous(
    name    = "Evolutionary time (years)",
    limits  = c(0, 15),
    breaks  = seq(0, 15, by = 1),
    sec.axis = sec_axis(
      trans = ~ . / generation_time_years,
      name  = "Transmission generations"
    )
  ) +
  scale_y_continuous(name = "Spatial distance (km)") +
  theme_bw(base_size = 14) +
  theme(
    panel.border       = element_blank(),
    panel.grid.major   = element_blank(),
    panel.grid.minor   = element_blank(),
    # bottom & left axes
    axis.line.x.bottom = element_line(color = "black"),
    axis.line.y.left   = element_line(color = "black"),
    # faint top axis line reconnecting the ticks
    axis.line.x.top    = element_line(color = "black", size = 0.3),
    axis.ticks.x.top   = element_line(color = "black"),
    axis.ticks.length  = unit(4, "pt"),
    axis.title.x.top   = element_text(margin = margin(b = 6), size = 12),
    axis.text.x.top    = element_text(size = 10),
    # no main title
    plot.title         = element_blank()
  )

all_serotype_evol_time_fixed


```






```{r}
library(dplyr)
library(boot)

early_data <- all_data_df %>%
  filter(!is.na(patristic_dist),
         !is.na(spatial_dist),
         patristic_dist <= 3)

slope_fun <- function(df, i) {
  d <- df[i, ]
  coef(lm(spatial_dist ~ patristic_dist, data = d))[2]
}

set.seed(42)
boot_slope <- boot(data      = early_data,
                   statistic = slope_fun,
                   R         = 1000)

slope_ci <- boot.ci(boot_slope, type = "perc")$percent[4:5]

slope_est <- coef(lm(spatial_dist ~ patristic_dist, data = early_data))[2]

spearman_test <- cor.test(early_data$patristic_dist,
                          early_data$spatial_dist,
                          method = "spearman",
                          exact  = FALSE)

cat(sprintf("Slope estimate: %.3f km/year\n", slope_est))
cat(sprintf("95%% CI for slope: [%.3f, %.3f] km/year\n",
            slope_ci[1], slope_ci[2]))
cat(sprintf("Spearman rho: %.3f\n", spearman_test$estimate))
cat(sprintf("95%% CI for rho: [%.3f, %.3f]\n",
            spearman_test$conf.int[1],
            spearman_test$conf.int[2]))
cat(sprintf("Spearman p‐value: %.3g\n", spearman_test$p.value))

```
#Sample size for bins 
```{r}
all_serotypes_binning_1G_fixed %>%
  filter(midpoint %in% c(0.5, 1.5, 2.5)) %>%
  select(midpoint, n)
```




#Individual serotype Bins 
```{r}
d1_binning_1G_fixed <- fc_1G_binning_fixed_median(d1_pairs_df) 
d2_binning_1G_fixed <- fc_1G_binning_fixed_median(d2_pairs_df) 
d3_binning_1G_fixed <- fc_1G_binning_fixed_median(d3_pairs_df) 
d4_binning_1G_fixed <- fc_1G_binning_fixed_median(d4_pairs_df) 
```

#Load Medians 
```{r}
d1_median <- d1_pairs_df %>% 
  summarise(median_spatial = median(spatial_dist, na.rm = TRUE)) %>% 
  pull(median_spatial)

d2_median <- d2_pairs_df %>% 
  summarise(median_spatial = median(spatial_dist, na.rm = TRUE)) %>% 
  pull(median_spatial)

d3_median <- d3_pairs_df %>% 
  summarise(median_spatial = median(spatial_dist, na.rm = TRUE)) %>% 
  pull(median_spatial)

d4_median <- d4_pairs_df %>% 
  summarise(median_spatial = median(spatial_dist, na.rm = TRUE)) %>% 
  pull(median_spatial)
```

#DENV 1 
```{r}
d1_evol_time <- ggplot(d1_binning_1G_fixed, aes(x = midpoint, y = spatial_median)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
              fill  = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = d1_median,
             linetype = "dashed",
             color    = "black") +
  scale_x_continuous(
    name    = "Evolutionary time (years)",
    limits  = c(0, 15),
    breaks  = seq(0, 15, by = 1),
    sec.axis = sec_axis( 
      trans = ~ . / generation_time_years,
      name  = "Transmission generations (No.)"
    )
  ) +
  theme_minimal() +
  labs(
    title = "DENV 1",
    y     = "Spatial distance (km)"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top  = element_text(size = 10),
    axis.line.x.top  = element_line(color = "black")
  )

d1_evol_time
```

#DENV 2
```{r}
d2_evol_time <- ggplot(d2_binning_1G_fixed, aes(x = midpoint, y = spatial_median)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
              fill  = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = d2_median,
             linetype = "dashed",
             color    = "black") +
  scale_x_continuous(
    name    = "Evolutionary time (years)",
    limits  = c(0, 15),
    breaks  = seq(0, 15, by = 1),
    sec.axis = sec_axis( 
      trans = ~ . / generation_time_years,
      name  = "Transmission generations (No.)"
    )
  ) +
  theme_minimal() +
  labs(
    title = "DENV 2",
    y     = "Spatial distance (km)"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top  = element_text(size = 10),
    axis.line.x.top  = element_line(color = "black")
  )

d2_evol_time
```

#DENV 3
```{r}
d3_evol_time <- ggplot(d3_binning_1G_fixed, aes(x = midpoint, y = spatial_median)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
              fill  = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = d3_median,
             linetype = "dashed",
             color    = "black") +
  scale_x_continuous(
    name    = "Evolutionary time (years)",
    limits  = c(0, 15),
    breaks  = seq(0, 15, by = 1),
    sec.axis = sec_axis( 
      trans = ~ . / generation_time_years,
      name  = "Transmission generations (No.)"
    )
  ) +
  theme_minimal() +
  labs(
    title = "DENV 3",
    y     = "Spatial distance (km)"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top  = element_text(size = 10),
    axis.line.x.top  = element_line(color = "black")
  )

d3_evol_time
```

#DENV 4
```{r}
d4_evol_time <- ggplot(d4_binning_1G_fixed, aes(x = midpoint, y = spatial_median)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
              fill  = "lightblue", alpha = 0.4) +
  geom_point(color = "blue", size = 1.5) +
  geom_line(color = "blue") +
  geom_hline(yintercept = d4_median,
             linetype = "dashed",
             color    = "black") +
  scale_x_continuous(
    name    = "Evolutionary time (years)",
    limits  = c(0, 10),
    breaks  = seq(0, 10, by = 1),
    sec.axis = sec_axis( 
      trans = ~ . / generation_time_years,
      name  = "Transmission generations (No.)"
    )
  ) +
  theme_minimal() +
  labs(
    title = "DENV 4",
    y     = "Spatial distance (km)"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top  = element_text(size = 10),
    axis.line.x.top  = element_line(color = "black")
  )

d4_evol_time
```

#All serotypes faceted separately 
```{r}
d3_binning_1G_fixed <- d3_binning_1G_fixed %>%
  mutate(serotype = "DENV-3")

d4_binning_1G_fixed <- d4_binning_1G_fixed %>%
  mutate(serotype = "DENV-4")

d1_binning_1G_fixed <- d1_binning_1G_fixed %>%
  mutate(serotype = "DENV-1")

d2_binning_1G_fixed <- d2_binning_1G_fixed %>%
  mutate(serotype = "DENV-2")

combined_binned_df <- bind_rows(d3_binning_1G_fixed, d4_binning_1G_fixed, d1_binning_1G_fixed, d2_binning_1G_fixed)

combined_overall_median <- bind_rows(
  d1_pairs_df,
  d2_pairs_df,
  d3_pairs_df,
  d4_pairs_df
) %>%
  pull(spatial_dist) %>%
  median(na.rm = TRUE)
```


```{r}
combined_evol_time <- ggplot(combined_binned_df, aes(x = midpoint, y = spatial_median, color = serotype, fill = serotype)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci), alpha = 0.3, color = NA) +
  geom_line(size = 1) +
  geom_point(size = 1.5) +
  geom_hline(yintercept = combined_overall_median, linetype = "dashed", color = "black") +
  
  scale_color_manual(values = c("DENV-3" = "blue", "DENV-4" = "darkred", "DENV-1" = "darkgreen", "DENV-2" = "purple")) +
  scale_fill_manual(values  = c("DENV-3" = "blue", "DENV-4" = "darkred", "DENV-1" = "darkgreen", "DENV-2" = "purple")) +
  
  scale_x_continuous(
    name    = "Evolutionary time (years)",
    limits  = c(0, 10),            
    breaks  = seq(0, 10, by = 1),    
    sec.axis = sec_axis(
      trans = ~ . / generation_time_years,
      name  = "Transmission generations (No.)"
    )
  ) +
  
  theme_minimal(base_size = 14) +
  theme(
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top  = element_text(size = 10),
    axis.line.x.top  = element_line(color = "black"),
    legend.title     = element_blank(),
    legend.position  = "top"
  ) +
  
  labs(
    y = "Spatial distance (km)"
  )

print(combined_evol_time)

```

##Not enough sample size for per period/serotype time period plots: complete period divide plots for all serotype comparisons 

#Binning for Plot 1G for All Serotype Time Comparison 
```{r}
period1_df <- period_lookup_df %>% 
  filter(period == "1994-2006") %>% 
  fc_1G_binning_fixed_median() %>% 
  mutate(period = "1994-2006")

period2_df <- period_lookup_df %>% 
  filter(period == "2007-2019") %>% 
  fc_1G_binning_fixed_median() %>% 
  mutate(period = "2007-2019")

period_cumulative_df <- bind_rows(period1_df, period2_df)

overall_median        <- bind_rows(d1_pairs_df, d2_pairs_df, d3_pairs_df, d4_pairs_df) %>% 
                         pull(spatial_dist) %>% 
                         median(na.rm = TRUE)
generation_time_years <- 20 / 365
```

#Plot 1G for All Serotype Time Comparison 
```{r}
all_sero_period_evol_time <- ggplot(period_cumulative_df, 
    aes(x = midpoint, y = spatial_median, color = period, fill = period)) +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
              alpha = 0.4, color = NA) +
  geom_point(size = 1.5) +
  geom_line(size = 1) +
  geom_hline(yintercept = overall_median,
             linetype = "dashed",
             color    = "black") +
  scale_x_continuous(
    name = "Evolutionary time (years)",
    sec.axis = sec_axis(
      trans = ~ . / generation_time_years,
      name  = "Transmission generations (No.)"
    )
  ) +
  scale_color_manual(values = c("1994-2006" = "blue",
                                "2007-2019" = "darkred")) +
  scale_fill_manual(values = c("1994-2006" = "blue",
                               "2007-2019" = "darkred")) +
  theme_minimal(base_size = 14) +
  theme(
    axis.title.x.top = element_text(margin = margin(b = 10), size = 12),
    axis.text.x.top  = element_text(size = 10),
    axis.line.x.top  = element_line(color = "black"),
    legend.title     = element_blank(),
    legend.position  = "top"
  ) +
  labs(y = "Spatial distance (km)")

all_sero_period_evol_time

#ggsave("1G_all_period.pdf", plot = all_sero_period_evol_time, device = "pdf", width = 8, height = 6)
```
```



