---
title: "matrices_kpp"
author: 
date: "`r Sys.Date()`"
output: html_document
---
#Load Packages
```{r}
library(librarian)
shelf(sf,
      data.table,
      ape,
      ggtree,
      phangorn,
      TreeTools,
      dplyr,
      tidyr,
      tibble, 
      readxl,
      stringr,
      tidyverse,
      BiocManager,
      treeio)
```

#MetaData Loading 
```{r}
#metadata
kpp_data <- read.csv("data/raw/kpp_data.csv", stringsAsFactors = FALSE)

#MAPPING DATA FOR NAMES
mapping_data_1 <- read_excel("data/raw/new_kpp_mapping.xlsx", sheet = 1, col_names = TRUE)
mapping_data_2 <- read_excel(
  path       = "data/raw/kpp_packing_list.xlsx",
  sheet      = 1,
  skip       = 1,        
  col_names  = TRUE)
mapping_data_list <- list(mapping_data_1,mapping_data_2)
```

#Standardize kpp_data timing 
```{r}
#make the timestamps and the dates standardized 
kpp_data$admitdate <- as.POSIXct(kpp_data$admitdate, 
                                 format = "%Y-%m-%dT%H:%M:%S", 
                                 tz = "UTC")

kpp_data$admitdate <- as.Date(kpp_data$admitdate, format="%Y-%m-%d")
```

#Make data subsets for computational efficiency
```{r}
# Filter and save for DEN1
den1_data <- kpp_data %>% filter(DENV.PCR == "DEN1")

# Filter and save for DEN2
den2_data <- kpp_data %>% filter(DENV.PCR == "DEN2")

# Filter and save for DEN3
den3_data <- kpp_data %>% filter(DENV.PCR == "DEN3")

# Filter and save for DEN4 (if it exists)
den4_data <- kpp_data %>% filter(DENV.PCR == "DEN4")
```

##HISTOGRAM TIME 
```{r}
#D1
hist(den1_data$study.year,
     main = "DENV 1 Distribution of Study Year",
     xlab = "Year",
     breaks = 30)

median(den1_data$study.year) #2008

#D2
hist(den2_data$study.year,
     main = "DENV 2 Distribution of Study Year",
     xlab = "Year",
     breaks = 30)

median(den2_data$study.year) #2009

#D3
hist(den3_data$study.year,
     main = "DENV 3 Distribution of Study Year",
     xlab = "Year",
     breaks = 30)

median(den3_data$study.year) #2013

#D4
hist(den4_data$study.year,
     main = "DENV 4 Distribution of Study Year",
     xlab = "Year",
     breaks = 30)

median(den4_data$study.year) #2015
```


#DISTANCE
#Source function for calculating euclidean distance
```{r}
source("script/clean/helper_scripts/helper_euclidean_distance.R")
```

#Calculate euclidean distance for all the serotypes 
```{r}
d1_distance <- fc_euclidean_dist(den1_data)
d2_distance <- fc_euclidean_dist(den2_data)
d3_distance <- fc_euclidean_dist(den3_data)
d4_distance <- fc_euclidean_dist(den4_data)
```

#TIME

#Source function for calculating time difference 
```{r}
source("script/clean/helper_scripts/helper_time_difference.R")
```

#Calculate time difference in days for all the serotypes 
```{r}
d1_time <- fc_time_difference(den1_data)
d2_time <- fc_time_difference(den2_data)
d3_time <- fc_time_difference(den3_data)
d4_time <- fc_time_difference(den4_data)
```


#PATRISTIC DISTANCES

#Load trees
```{r}
#TREES REPLACED
d3_kpp_thailand_consensus <- read.beast("data/beast_output/d3_st/d3_st_final.tree")
d4_kpp_thailand_consensus <- read.beast("data/beast_output/d4_st/d4_st_final.tree")
d1_kpp_thailand_consensus <- read.beast("data/beast_output/d1_st/d1_st_final.tree")
d2_kpp_thailand_consensus <- read.beast("data/beast_output/d2_st/d2_st_final.tree")
```

#Source function for calculating cophenetic partistic distance
```{r}
source("script/clean/helper_scripts/helper_patristic_cophyletic.R")
```

#Load the patristic data frames 
#With consensus trees 
```{r}
d4_patristic_cophen <- fc_patristic_consensus(d4_kpp_thailand_consensus, mapping_data_1)
d3_patristic_cophen <- fc_patristic_consensus(d3_kpp_thailand_consensus, mapping_data_1)
d1_patristic_cophen <- fc_patristic_consensus(d1_kpp_thailand_consensus, mapping_data_1)
d2_patristic_cophen <- fc_patristic_consensus(d2_kpp_thailand_consensus, mapping_data_1)
```






