---
title: "mrca_relative_risk"
author: 
date: "`r Sys.Date()`"
output: html_document
---
#Load packages 
```{r}
library(librarian)
shelf(dplyr)
shelf(tidyr)
shelf(ggplot2)  
```

#Load helper function 
```{r}
source("script/clean/helper_scripts/helper_mrca_risk.R")
```

#Bin the data frame by the MRCA and spatial groups 

#purpose of function = takes four counts
#Continuity‐corrected proportions
```{r}
df_binned3 <- all_data_df %>%
  filter(!is.na(spatial_dist), !is.na(mrca_age)) %>%
  mutate(
    mrca_bin = cut(
      mrca_age,
      breaks = c(0.0, 0.5, 2.0, 5.0, 10.0),
      labels = c("<0.5", "0.5–2", "2–5", "5–10"),
      right = TRUE,
      include.lowest = TRUE
    ),
    spatial_bin = cut(
      spatial_dist,
      breaks = c(0.0, 0.5, 10.0, Inf),
      labels = c("<0.5 km", "0.5–10 km", ">10 km"),
      right = TRUE,
      include.lowest = TRUE
    )
  ) %>%
  filter(!is.na(mrca_bin), !is.na(spatial_bin))

print(table(df_binned3$spatial_bin, df_binned3$mrca_bin))

as.data.frame(table(df_binned3$spatial_bin, df_binned3$mrca_bin)) %>%
  rename(spatial_bin = Var1, mrca_bin = Var2, n = Freq) %>%
  arrange(spatial_bin, mrca_bin) %>%
  print()

```

#Make a big count data frame 
```{r}
count_df3 <- df_binned3 %>%
  group_by(spatial_bin, mrca_bin) %>%
  summarise(n = n(), .groups = "drop")

all_grid3 <- expand.grid(
  spatial_bin = levels(df_binned3$spatial_bin),
  mrca_bin    = levels(df_binned3$mrca_bin),
  stringsAsFactors = FALSE
) %>% as_tibble()

count_df_full3 <- all_grid3 %>%
  left_join(count_df3, by = c("spatial_bin", "mrca_bin")) %>%
  mutate(n = ifelse(is.na(n), 0L, n))

count_df_full3 %>%
  arrange(spatial_bin, mrca_bin) %>%
  print()
```

#Make a total by spatial distance bin
#Make a reference >10 km bin
```{r}
totals_by_spatial3 <- count_df_full3 %>%
  group_by(spatial_bin) %>%
  summarise(total_in_spatial = sum(n), .groups = "drop")

ref_total3 <- totals_by_spatial3 %>%
  filter(spatial_bin == ">10 km") %>%
  pull(total_in_spatial)

if (length(ref_total3) == 0 || ref_total3 == 0) {
  stop("ERROR: There are no >10 km pairs => no reference data. Aborting.")
}
```

#Make a big count table 
```{r}
rr_prep_base3 <- count_df_full3 %>%
  left_join(totals_by_spatial3, by = "spatial_bin") %>%
  select(spatial_bin, mrca_bin, n, total_in_spatial)

ref_counts3 <- rr_prep_base3 %>%
  filter(spatial_bin == ">10 km") %>%
  select(mrca_bin, n_ref = n)

rr_prep3 <- rr_prep_base3 %>%
  left_join(ref_counts3, by = "mrca_bin") %>%
  mutate(
    n_ref    = ifelse(is.na(n_ref), 0L, n_ref),
    total_ref = ref_total3
  ) %>%
  mutate(
    p_s   = ifelse(total_in_spatial == 0, 0, n / total_in_spatial),
    p_ref = ifelse(total_ref         == 0, 0, n_ref / total_ref)
  )
rr_prep3 %>%
  arrange(spatial_bin, mrca_bin) %>%
  print()
```

```{r}
rr_results3 <- rr_prep3 %>%
  rowwise() %>%
  mutate(
    ci_vals = list(compute_rr_ci(n, total_in_spatial, n_ref, total_ref))
  ) %>%
  unnest_wider(ci_vals) %>%
  ungroup()

final_rr3 <- rr_results3 %>%
  select(
    spatial_bin,
    mrca_bin,
    n_S_W       = n,
    total_S     = total_in_spatial,
    n_ref_W     = n_ref,
    total_ref,
    p_s,
    p_ref,
    RR,       
    CI_lower,  
    CI_upper  
  ) %>%
  arrange(spatial_bin, mrca_bin)

rr_unadj3 <- rr_prep3 %>%
  rowwise() %>%
  mutate(
    ci_unadj_vals = list(compute_rr_ci(n, total_in_spatial, n_ref, total_ref))
  ) %>%
  unnest_wider(ci_unadj_vals) %>%
  ungroup() %>%
  rename(
    RR_unadj       = RR,
    CI_lower_unadj = CI_lower,
    CI_upper_unadj = CI_upper
  ) %>%
  select(spatial_bin, mrca_bin, RR_unadj, CI_lower_unadj, CI_upper_unadj)

final_rr_with_unadj3 <- final_rr3 %>%
  left_join(rr_unadj3, by = c("spatial_bin", "mrca_bin"))
```

#graph of the plot: UNADJUSTED
```{r}
plot_unadj3_df <- final_rr_with_unadj3 %>%
  mutate(
    spatial_bin = factor(
      spatial_bin,
      levels = c("<0.5 km", "0.5–10 km", ">10 km"),
      labels = c("<0.5", "0.5–10", ">10")
    ),
    facet_label = factor(
      mrca_bin,
      levels = c("<0.5", "0.5–2", "2–5", "5–10"),
      labels = c(
        "(A) MRCA < 6 months",
        "(B) MRCA 6 months–2 years",
        "(C) MRCA 2–5 years",
        "(D) MRCA 5–10 years"
      )
    ),
    is_ref = (as.character(spatial_bin) == ">10")
  )

ggplot(plot_unadj3_df, aes(x = spatial_bin, y = RR_unadj, color = is_ref)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50") +
  geom_errorbar(
    aes(ymin = CI_lower_unadj, ymax = CI_upper_unadj),
    width = 0.25, size = 0.6, na.rm = TRUE
  ) +
  geom_point(shape = 15, size = 3, na.rm = TRUE) +
  scale_y_log10(
    limits = c(0.01, 100),
    breaks = c(0.01, 0.1, 1, 10, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_color_manual(
    values = c("FALSE" = "black", "TRUE" = "red"),
    guide = FALSE
  ) +
  facet_wrap(~ facet_label, nrow = 2, ncol = 2, scales = "fixed") +
  labs(
    x = "Spatial distance (km)", 
    y = "Relative Risk"
  ) +
  theme_bw(base_size = 14) +
  theme(
    panel.grid.minor    = element_blank(),
    panel.grid.major    = element_line(color = "gray85"),
    axis.text.x         = element_text(angle = 0, hjust = 0.5, vjust = 1),
    axis.title.x        = element_text(margin = margin(t = 8)), 
    strip.text          = element_text(face = "bold", size = 12, hjust = 0),
    plot.margin         = margin(8, 8, 8, 8)
  )

```

```{r}
ggplot(plot_unadj3_df, aes(x = spatial_bin, y = RR_unadj, color = is_ref)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50") +
  geom_errorbar(
    aes(ymin = CI_lower_unadj, ymax = CI_upper_unadj),
    width = 0.25, size = 0.6, na.rm = TRUE
  ) +
  geom_point(shape = 15, size = 3, na.rm = TRUE) +
  scale_y_log10(
    limits = c(0.01, 100),
    breaks = c(0.01, 0.1, 1, 10, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_color_manual(
    values = c("FALSE" = "black", "TRUE" = "red"),
    guide = FALSE
  ) +
  facet_wrap(~ facet_label, nrow = 2, ncol = 2, scales = "fixed") +
  labs(
    x = "Spatial distance (km)", 
    y = "Relative Risk"
  ) +
  theme_bw(base_size = 14) +
  theme(
    panel.grid.major = element_blank(),  # remove major gridlines
    panel.grid.minor = element_blank(),  # remove minor gridlines
    axis.text.x      = element_text(angle = 0, hjust = 0.5, vjust = 1),
    axis.title.x     = element_text(margin = margin(t = 8)), 
    strip.text       = element_text(face = "bold", size = 12, hjust = 0),
    plot.margin      = margin(8, 8, 8, 8)
  )

```


```{r}
library(dplyr)

rr_numbers <- final_rr_with_unadj3 %>%
  arrange(mrca_bin, spatial_bin) %>%
  select(
    MRCA       = mrca_bin,
    Spatial    = spatial_bin,
    RR_adj     = RR,
    CI_low     = CI_lower,
    CI_high    = CI_upper
  )

print(rr_numbers)
```



#Make bins with rolling windows instead
```{r}
window_starts <- seq(0, 9, by = 1)

time_to_mixing_list <- lapply(window_starts, function(t0) {
  df_win <- df_binned3 %>%
    filter(mrca_age >= t0, mrca_age < (t0 + 1))
  
  count_win <- df_win %>%
    group_by(spatial_bin) %>%
    summarise(n_SW = n(), .groups = "drop")
  
  merged <- totals_by_spatial3 %>%
    left_join(count_win, by = "spatial_bin") %>%
    mutate(
      n_SW = ifelse(is.na(n_SW), 0L, n_SW),     
      n_S  = total_in_spatial                  
    )
  
  n_ref_W <- count_win %>%
    filter(spatial_bin == ">10 km") %>%
    pull(n_SW)
  if (length(n_ref_W) == 0) {
    n_ref_W <- 0L
  }
  
  merged <- merged %>%
    mutate(
      n_ref_W   = n_ref_W,      
      total_ref = ref_total3     
    )
  
  merged <- merged %>%
    rowwise() %>%
    mutate(
      ci_out = list(compute_rr_ci(n_SW, n_S, n_ref_W, total_ref))
    ) %>%
    unnest_wider(ci_out) %>% 
    ungroup()
  
  merged <- merged %>%
    mutate(window_start = t0)
  
  return(merged)
})

df_time_mixing <- bind_rows(time_to_mixing_list)
```


#Plot the rolling window graph 
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

window_starts <- seq(0, 9, by = 1)

time_to_mixing_list <- lapply(window_starts, function(t0) {
  df_win <- df_binned3 %>%
    filter(mrca_age >= t0, mrca_age < (t0 + 1))
  
  count_win <- df_win %>%
    group_by(spatial_bin) %>%
    summarise(n_SW = n(), .groups = "drop")
  
  merged <- totals_by_spatial3 %>%
    left_join(count_win, by = "spatial_bin") %>%
    mutate(
      n_SW = ifelse(is.na(n_SW), 0L, n_SW),     
      n_S  = total_in_spatial                  
    )
  
  n_ref_W <- count_win %>%
    filter(spatial_bin == ">10 km") %>%
    pull(n_SW)
  if (length(n_ref_W) == 0) {
    n_ref_W <- 0L
  }
  
  merged <- merged %>%
    mutate(
      n_ref_W   = n_ref_W,      
      total_ref = ref_total3     
    )
  
  merged <- merged %>%
    rowwise() %>%
    mutate(
      ci_out = list(compute_rr_ci(n_SW, n_S, n_ref_W, total_ref))
    ) %>%
    unnest_wider(ci_out) %>% 
    ungroup()
  
  merged <- merged %>%
    mutate(
      window_start = t0,
      window_mid   = t0 + 0.5,
      spatial_lbl  = spatial_bin
    )
  
  return(merged)
})

df_time_mixing <- bind_rows(time_to_mixing_list)

ggplot(df_time_mixing, 
       aes(x = window_mid, y = RR, 
           color = spatial_lbl, linetype = spatial_lbl)) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", size = 0.5) +
  geom_errorbar(
    aes(ymin = CI_lower, ymax = CI_upper),
    width = 0.1, size = 0.6, alpha = 0.8, show.legend = FALSE
  ) +
  geom_line(size = 1.2) +
  geom_point(size = 2.5, shape = 16, stroke = 0.8, fill = "white") +
  scale_y_log10(
    limits = c(0.01, 100),
    breaks = c(0.01, 0.1, 1, 10, 100),
    labels = c("0.01", "0.1", "1", "10", "100")
  ) +
  scale_x_continuous(
    breaks = seq(0, 9, by = 1),
    labels = paste0(seq(0, 9, by = 1), "–", seq(1, 10, by = 1)),
    expand = expansion(add = c(0.2, 0.2))  
  ) +
  scale_color_manual(
    name = "Distance bin",
    values = c(
      "<0.5 km"   = "#1b9e77", 
      "0.5–10 km" = "#d95f02",  
      ">10 km"    = "#7570b3"   
    )
  ) +
  scale_linetype_manual(
    name = "Distance bin",
    values = c(
      "<0.5 km"   = "solid",  
      "0.5–10 km" = "dashed",   
      ">10 km"    = "dotted"     
    )
  ) +
  labs(
    x        = "MRCA (Years)",
    y        = "Relative Risk"
  ) +
  theme_bw(base_size = 14) +
  theme(
    plot.title = element_text(
      face = "bold", size = 16, margin = margin(b = 4),
      hjust = 0.5
    ),
    plot.subtitle = element_text(
      size = 12, face = "italic", margin = margin(b = 12),
      hjust = 0.5
    ),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "gray85", size = 0.5),
    axis.text.x  = element_text(size = 12, angle = 0, hjust = 0.5),
    axis.text.y  = element_text(size = 12),
    axis.title.x = element_text(size = 14, margin = margin(t = 8)),
    axis.title.y = element_text(size = 14, margin = margin(r = 8)),
    legend.position    = "top",
    legend.title       = element_text(size = 13, face = "bold"),
    legend.text        = element_text(size = 12),
    legend.key.width   = unit(2, "lines"),
    plot.margin        = margin(t = 8, r = 8, b = 8, l = 8)
  )

```

