---
title: "sex_comparison"
author: "Tanmayi Patharkar"
date: "`r Sys.Date()`"
output: html_document
---
```{r}
library(librarian)
shelf(dplyr)
shelf(binom)
shelf(tidyr)
shelf(ggplot2)
shelf(scales)
shelf(purrr)
shelf(tibble)
```

#Source 2A Binning Function
```{r}
source("script/clean/helper_scripts/helper_2A_binning.R")
```


#At each cumulative distance, how much more likely is a male case than a female case to share the same transmission chain


##NEW BINNING 

#MAKE SEX DF
```{r}
sex_pair_df <- all_data_df %>%
  dplyr::left_join(
    dplyr::select(kpp_data, studyno, admitdate, sex),
    by = c("sample1" = "studyno")
  ) %>%
  dplyr::rename(admit1 = admitdate, sex1 = sex) %>%
  dplyr::mutate(admit1 = as.Date(admit1)) %>%
  dplyr::left_join(
    dplyr::select(kpp_data, studyno, admitdate, sex),
    by = c("sample2" = "studyno")
  ) %>%
  dplyr::rename(admit2 = admitdate, sex2 = sex) %>%
  dplyr::mutate(admit2 = as.Date(admit2))

```


#ORDERED SEX DF 
```{r}
ordered_sex_df <- sex_pair_df %>%
  dplyr::mutate(first1 = admit1 <= admit2) %>%
  dplyr::mutate(
    sample1 = if_else(first1, sample1, sample2),
    sample2 = if_else(first1, sample2, sample1),
    admit1  = if_else(first1, admit1,  admit2),
    admit2  = if_else(first1, admit2,  admit1),
    sex1    = if_else(first1, sex1,    sex2),
    sex2    = if_else(first1, sex2,    sex1)
  ) %>%
  dplyr::select(-first1)

ordered_sex_df <- ordered_sex_df %>%
  dplyr::mutate(
    sex1 = factor(sex1, levels = c("F", "M"))
  )

ordered_sex_df <- ordered_sex_df %>%
  filter(!is.na(spatial_dist)) %>%
  mutate(
    log_dist = log(spatial_dist + 1e-6) 
  )

```


#RUN BINNING FUNCTION 
```{r}
female_df <- ordered_sex_df %>%
  filter(sex1 == "F") %>%
  fc_2A_binning()

male_df <- ordered_sex_df %>%
  filter(sex1 == "M") %>%
  fc_2A_binning()

panelE_df <- dplyr::inner_join(
  female_df, male_df,
  by     = "bin_cutoff",
  suffix = c("_F", "_M")
) %>%
  dplyr::mutate(
    diff       = prop_same_chain_M - prop_same_chain_F,  
    se_diff    = sqrt(se_F^2 + se_M^2),
    lower_diff = diff - 1.96 * se_diff,
    upper_diff = diff + 1.96 * se_diff
  )
```

###PLOT
```{r}
#plot it 
ggplot(panelE_df, aes(x = bin_cutoff, y = diff)) +
  geom_ribbon(aes(ymin = lower_diff, ymax = upper_diff),
              fill = "#7a0177", alpha = 0.3) +
  geom_line(size = 1.2, color = "#7a0177") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +

  scale_y_continuous(
    name   = "Difference in prop. same chain\n(M vs F) [Female Baseline]",
    limits = c(min(panelE_df$lower_diff, -0.01),
               max(panelE_df$upper_diff)),
    breaks = seq(-0.01, 0.03, by = 0.005),
    expand = c(0, 0)
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, by = 5),
    labels = scales::number_format(accuracy = 1)
  ) +

  theme_minimal(base_size = 14) +
  theme(
    panel.grid   = element_blank(),
    axis.line    = element_line(size = 0.6, color = "black"),
    axis.ticks   = element_line(size = 0.6, color = "black"),
    axis.title   = element_text(size = 14, face = "plain"),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.text    = element_text(size = 12)
  )
```

#GLM 
```{r}
glm_sex_dist <- glm(
  same_chain    ~ sex1 * spatial_dist,
  data          = ordered_sex_df,      
  family        = binomial(link = "logit")
)

summary(glm_sex_dist)

library(broom)
or_table <- broom::tidy(glm_sex_dist, exponentiate=TRUE, conf.int=TRUE) %>%
  dplyr::select(term, OR = estimate, CI_low = conf.low, CI_high = conf.high, p.value) %>%
  dplyr::mutate_at(dplyr::vars(OR:CI_high, p.value), ~ round(., 3))
print(or_table)
```


#########LOG AND NEWEST MODEL 
```{r}
library(dplyr)
library(broom)
library(multiwayvcov)
library(lmtest)

ordered_sex_df <- ordered_sex_df %>%
  filter(!is.na(spatial_dist)) %>%
  mutate(
    log_dist = log(spatial_dist + 1e-6)  # tiny constant to avoid log(0)
  )

glm_sex_log <- glm(
  same_chain ~ sex1 * log_dist,
  data   = ordered_sex_df,
  family = binomial(link = "logit")
)

vcov_sex_cl <- cluster.vcov(
  glm_sex_log,
  cluster = list(
    ordered_sex_df$sample1,
    ordered_sex_df$sample2
  )
)

age_or_table_sex_log_robust <- broom::tidy(
  glm_sex_log,
  conf.int     = TRUE,
  exponentiate = TRUE,
  vcov.        = vcov_sex_cl
) %>%
  mutate(
    OR       = sprintf("%.3f", estimate),
    `95% CI` = sprintf("[%.3f–%.3f]", conf.low, conf.high),
    `P-value`= ifelse(p.value < 0.001, "<0.001", sprintf("%.3f", p.value))
  ) %>%
  select(
    Term    = term,
    OR,
    `95% CI`,
    `P-value`
  )

age_or_table_sex_log_robust
```


```{r}
library(pROC)

mf <- model.frame(glm_sex_log)

response <- model.response(mf)                   
pred_prob <- predict(glm_sex_log, newdata = mf, type = "response")


roc_obj <- roc(response, pred_prob)
auc_val <- auc(roc_obj)
cat("AUC =", round(auc_val, 3), "\n")

brier <- mean((pred_prob - response)^2)
cat("Brier score =", round(brier, 4), "\n")
```












###############################

#F-F BINNING 
```{r}
ff_binned <- sex_data_df %>%
  filter(sex1 == "F", sex2 == "F") %>%   
  fc_2A_binning() %>%                   
  rename(
    prop_FF   = prop_same_chain,         
    lower_FF  = lower,
    upper_FF  = upper
  )
```


#F-F pairs graphed 
```{r}
ggplot(ff_binned, aes(x = bin_cutoff, y = prop_FF)) +
  geom_ribbon(
    aes(ymin = lower_FF, ymax = upper_FF),
    fill  = "#7a0177", 
    alpha = 0.3
  ) +
  geom_line(size = 1.2, color = "#7a0177") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  scale_y_continuous(
    name   = "P(same chain)\n(F–F pairs only)",
    limits = c(min(ff_binned$lower_FF, -0.005),   
               max(ff_binned$upper_FF, 0.05)),    
    breaks = seq(-0.005, 0.05, by = 0.005),       
    labels = scales::number_format(accuracy = 0.005),
    expand = c(0, 0)
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, by = 5),
    labels = scales::number_format(accuracy = 1),
    expand = c(0, 0)
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid   = element_blank(),
    axis.line    = element_line(size = 0.6, color = "black"),
    axis.ticks   = element_line(size = 0.6, color = "black"),
    axis.title   = element_text(size = 14, face = "bold"),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.text    = element_text(size = 12)
  )
```

##F–F cases closer together are more likely than random to be in the same transmission chain.

```{r}
mm_binned <- sex_data_df %>%
  filter(sex1 == "M", sex2 == "M") %>%  
  fc_2A_binning() %>%                    
  rename(
    prop_mm   = prop_same_chain,        
    lower_mm = lower,
    upper_mm  = upper
  )
```


```{r}
ggplot(mm_binned, aes(x = bin_cutoff, y = prop_mm)) +
  geom_ribbon(
    aes(ymin = lower_mm, ymax = upper_mm),
    fill  = "#7a0177", 
    alpha = 0.3
  ) +
  geom_line(size = 1.2, color = "#7a0177") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  scale_y_continuous(
    name   = "P(same chain)\n(M-M pairs only)",
    limits = c(min(mm_binned$lower_mm, -0.005),   
               max(mm_binned$upper_mm, 0.05)),     
    breaks = seq(-0.005, 0.05, by = 0.005),       
    labels = scales::number_format(accuracy = 0.005),
    expand = c(0, 0)
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, by = 5),
    labels = scales::number_format(accuracy = 1),
    expand = c(0, 0)
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid   = element_blank(),
    axis.line    = element_line(size = 0.6, color = "black"),
    axis.ticks   = element_line(size = 0.6, color = "black"),
    axis.title   = element_text(size = 14, face = "bold"),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.text    = element_text(size = 12)
  )
```

##M-M cases closer together are more likely than random to be in the same transmission chain, more than females 

#M and F comparison 

```{r}
comparison_df <- full_join(mm_binned, ff_binned, by = "bin_cutoff") %>%
  mutate(
    diff_prop  = prop_mm - prop_FF,
    se_mm = (upper_mm - lower_mm) / (2 * 1.96),
    se_ff = (upper_FF - lower_FF) / (2 * 1.96),
    se_diff = sqrt(se_mm^2 + se_ff^2),
    lower_diff = diff_prop - 1.96 * se_diff,
    upper_diff = diff_prop + 1.96 * se_diff
  )

ggplot(comparison_df, aes(x = bin_cutoff, y = diff_prop)) +
  geom_ribbon(aes(ymin = lower_diff, ymax = upper_diff), fill = "#0570b0", alpha = 0.3) +
  geom_line(size = 1.2, color = "#0570b0") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
  scale_y_continuous(
    name   = "Difference in P(same chain)\n(M–M minus F–F)",
    breaks = seq(-0.03, 0.04, by = 0.005),
    expand = c(0, 0)
  ) +
  scale_x_continuous(
    name   = "Distance (km)",
    limits = c(0, 30),
    breaks = seq(0, 30, 5),
    expand = c(0, 0)
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid   = element_blank(),
    axis.line    = element_line(size = 0.6, color = "black"),
    axis.ticks   = element_line(size = 0.6, color = "black"),
    axis.title   = element_text(size = 14, face = "bold"),
    axis.text.x  = element_text(angle = 45, hjust = 1),
    axis.text    = element_text(size = 12)
  )

```

#males are more likely than females to transmit locally
#M–M transmission is more spatially localized than F–F.






####ASSORTATIVE ANALYSIS FOR SEX 
```{r}
shelf(dplyr)
shelf(speedglm)
shelf(tibble)

set.seed(2025)
df_sex <- sex_pair_df %>%
  filter(!is.na(spatial_dist)) %>%
  mutate(
    log_dist     = log(spatial_dist + 1e-6),
    same_sex_num = as.integer(sex1 == sex2),
    serotype     = factor(serotype)
  )

fmla_sex <- same_chain ~ same_sex_num * (log_dist + serotype)

two_way_boot <- function(data, formula, cl1, cl2, B = 500) {
  ids1 <- unique(data[[cl1]])
  ids2 <- unique(data[[cl2]])
  n1   <- length(ids1)
  n2   <- length(ids2)
  
  base_fit <- speedglm(formula, data, family=binomial(link="logit"), accel=TRUE)
  coef_names <- names(coef(base_fit))
  p          <- length(coef_names)
  
  boots <- matrix(NA, nrow = B, ncol = p, dimnames = list(NULL, coef_names))
  
  for (b in seq_len(B)) {
    draw1 <- sample(ids1, size = n1, replace = TRUE)
    draw2 <- sample(ids2, size = n2, replace = TRUE)
    
    boot_df <- data %>%
      filter( .data[[cl1]] %in% draw1,
              .data[[cl2]] %in% draw2 )
    
    fit_b <- speedglm(
      formula = formula,
      data    = boot_df,
      family  = binomial(link = "logit"),
      accel   = TRUE
    )
    
    boots[b, ] <- coef(fit_b)
  }
  
  ci_log   <- apply(boots, 2, quantile, probs = c(0.025, 0.975))
  orig_log <- coef(base_fit)
  
  tibble(
    Term     = coef_names,
    OR       = sprintf("%.3f", exp(orig_log)),
    `95% CI` = sprintf("[%.3f–%.3f]",
                       exp(ci_log[1, ]),
                       exp(ci_log[2, ]))
  )
}

reg_table_sex <- two_way_boot(df_sex, fmla_sex, "sample1", "sample2", B = 500)
print(reg_table_sex)

```

```{r}
library(dplyr)
library(tidyr)

reg_table_sex_sig <- reg_table_sex %>%
  # 1) strip brackets from the CI
  mutate(CI_clean = gsub("^\\[|\\]$", "", `95% CI`)) %>%
  # 2) split into numeric lo / hi
  separate(CI_clean, into = c("lo", "hi"), sep = "–", convert = TRUE) %>%
  # 3) define significance
  mutate(significant = if_else(lo > 1 | hi < 1, "yes", "no")) %>%
  # 4) pick exactly these four columns
  dplyr::select(Term, OR, `95% CI`, significant)

print(reg_table_sex_sig)

```

